---
title: "Accomplishments"
subtitle: "GWAS of the RDoC Cognitive Systems Domain: Modeling the Latent Genetic Architecture of Working Memory"
author:
  - name: "PI: Joey Trampush PhD"
date: January 1, 2021
output:
  komadown::komatufte:
    latex_engine: pdflatex
    citation_package: biblatex
    number_sections: true
    keep_md: true
    keep_tex: false
site: bookdown::bookdown_site
toc: false
highlight: yes
bibliography: komadown.bib
biblio-style: apa
link-citations: yes
colorlinks: yes
graphics: yes
fig-caption: true
self-contained: yes
fontpack-libertine: yes
fontpack-newtx: yes
automark: yes
classoption:
  - DIV=calc
  - headsepline=true
header:
  - pos: ro
    next: "GWAS of Working Memory"
caption:
  - labelfont=bf
  - labelsep=period
  - font=small
setkomafont:
  - element: descriptionlabel
    commands: \normalfont\scshape\bfseries
---

```{r, setup, include = FALSE, results='hide', warning=FALSE}
library(knitr)
knitr::knit_hooks$set(crop = knitr::hook_pdfcrop)
knitr::knit_hooks$set(optipng = knitr::hook_optipng)

# Fix subfigs for rmarkdown -> latex
if (identical(knitr:::pandoc_to(), 'latex'))
  knitr::knit_hooks$set(plot = knitr::hook_plot_tex)

knitr::opts_chunk$set(
  echo = FALSE,
  dev = c("CairoPNG", "CairoPDF"),
  dev.args = list(pointsize = 8),
  crop = TRUE,
  par = TRUE,
  fig.width = 5,
  fig.height = 3,
  dpi = 300,
  fig.pos = '!ht',
  root.dir = normalizePath('./'),
  cache = FALSE,
  cache.path = './cache/report-',
  # Code block options
  eval = TRUE,
  include = TRUE,
  message = FALSE,
  warning = FALSE,
  results = 'asis',
  error = TRUE
)

knitr::opts_knit$set(width = 50)

options(digits = 3, warnPartialMatchArgs = FALSE)
```

```{r}
# Source functions
source(file = here::here("R/utils.R"), local = knitr::knit_global())
```

```{r}
# Source functions
source(file = here::here("R/handout.R"), local = knitr::knit_global())
```

```{r}
# Source functions
source(file = here::here("R/html.R"), local = knitr::knit_global())
```

```{r}
knitr::knit_engines$set(marginfigure = function(options) {
        options$type = 'marginfigure'
        eng_block = knitr::knit_engines$get('block')
        eng_block(options)
    })
```

(ref:zplot) *z* = -1 (16th percentile), *z* = 0 (50th percentile), and *z* = 1
(84th percentile)

(ref:pct) A percentile score shows a person's standing relative to peers. For
example, a percentile score of 70 means Biggie performed as well as or better
than 70% of individuals the same age.

(ref:br) A base rate is the prevalence of score in the general population. If a
score is rare, defined as a base rate of less than 10%, it is considered
*clinically significant*, and that area of functioning is probably causing
problems in daily functioning.

(ref:vrb) *Verbal/language* refers to the ability to access and apply acquired
word knowledge, to verbalize meaningful concepts, to understand complex
multistep instructions, to think about verbal information, and to express
oneself using words.

(ref:spt) *Visuoperception, visuoconstruction, and visuospatial processing*
refer to abilities such as mentally visualizing how objects should look from
different angles, visualizing how to put objects together so that they fit
correctly, and being able to accurately and efficiently copy and/or reproduce
visual-spatial information onto paper.

(ref:att-exe) *Attentional* and *executive functions* underlie most, if not all,
domains of cognitive performance. These are behaviors and skills that allow
individuals to successfully carry-out instrumental and social activities,
academic work, engage with others effectively, problem solve, and successfully
interact with the environment to get needs met.

(ref:att) *Attention* is a basic function that underlies most, if not all,
domains of neurocognitive functioning. If an individual has limited attentional
capacity, an inability to sustain or divide attentional focus, or a reduced
ability to selectively attend to competing stimuli, then they would likely be
unable to interact with the environment to an adequate degree and, thus have
limited functioning in other, higher-order areas.

(ref:exe) Measures of *executive functioning* are used as proxies for behaviors
and skills that allow individuals to successfully carry-out instrumental and
social activities, such as planning, engaging with others effectively, problem
solving, and successfully interacting with the environment to get needs met.

(ref:wm) *Working memory* is the system responsible for the transient holding
and processing of information in conscious awareness just long enough to use it
for problem solving, and plays a critical role in reasoning, comprehension, and
learning.

(ref:ps) *Processing speed* refers to the rate at which the brain can
repetitively process familiar and unfamiliar information during relatively easy
tasks that almost everyone can do. Processing speed tests are good predictors of
performance on tasks that are so well learned that they are nearly automatic,
which can be critical to completing work efficiently.

(ref:mem) *Learning and memory* refer to the rate and ease with which new
information (e. g., facts, stories, lists, faces, names) can be encoded, stored,
and later recalled from long-term memory.

(ref:mtr) *Sensorimotor* tasks refer to the capacity to control hand movements
quickly, smoothly, and with adequate precision, which are required to engage in
activities such as writing and drawing.

(ref:soccog) *Social cognition* measures are used to gauge aspects of
social-emotional processing and social aptitude (sometimes referred to as
emotional intelligence), which are important facets of how individuals process
social information about other children, adults, groups, and social contexts.

(ref:cc) *Cognitive control* guides behavior by controlling what, when, and how
information is represented in the brain. Cognitive control is associated with a
wide range of processes and is not restricted to a particular cognitive domain.
Impairments in cognitive control functions are associated with specific deficits
in attention, memory, language comprehension and emotional processing.

(ref:adhd) *ADHD* is a neurodevelopmental disorder characterized by symptoms of
inattention (not being able to keep focus), hyperactivity (excess movement that
is not fitting to the setting) and impulsivity (hasty acts that occur in the
moment without thought). Many ADHD symptoms, such as high activity levels,
difficulty remaining still for long periods of time and limited attention spans,
are common to children and adolescents in general, and even many adults. The
difference in individuals with ADHD is that their hyperactivity and inattention
are noticeably greater than expected for their age and cause distress and/or
problems functioning at home, at school, at work, or with friends.

# Major activities

Mostly data tidying and imputation.

# Specific objectives

The field moves fast; tech improves.

# Significant results

-   including major findings, developments, or conclusions (both positive and
    negative)

# Key outcomes or other achievements

Include a discussion of stated goals not met. As the project progresses, the
emphasis in reporting in this section should shift from reporting activities to
reporting accomplishments. Goals are equivalent to specific aims. In the
response, emphasize the significance of the findings to the scientific field.
Include the approaches taken to ensure robust and unbiased results. For most NIH
awards the response should not exceed 2 pages.

## High-Performance Computing

Computational research often requires resources that exceed those of a personal
laptop or desktop computer. High-performance computing aggregates the resources
from individual computers (known as nodes) into a cluster that works together to
perform advanced, specialized computing jobs. The Center for Advanced Research
Computing launched its new high-performance computing cluster, Discovery, in
August 2020. The new cluster marks a significant upgrade to the CARC's
cyberinfrastructure, and the first step in a major, user-focused overhaul of the
program. The Discovery cluster consists of 2 shared login nodes and a total of
around 11,000 CPU cores in around 500 compute nodes. The typical compute node
has dual 8 to 16 core processors and resides on a 56 gigabit FDR InfiniBand
backbone. Nearly 100 GB of storage space is provided for each user to store
important code and configuration files in their home directories, and the
BeeGFS/ZFS parallel project file system has a capacity of 8.4 PB of usable
space, with a default storage quota of 10 TB per Principal Investigator across
their projects. A rapidly growing fleet of state-of-the art multicore compute
nodes is available for data-intensive research jobs, backed by two 800 TB,
high-throughput scratch file systems. A low-latency, high-bandwidth InfiniBand
network fabric facilitates intense research workloads on large data sets.
Discovery includes an array of scientific software packages, both licensed and
open source, for engineering, molecular simulation, and computational chemistry.
Researchers can also install software packages or develop their own code within
their project's allotted storage. Discovery is free to use for all USC faculty,
research staff, and graduate students (with the approval of their faculty
advisor). For information on how to set up your account and start using the
cluster, see the Accounts page. If you anticipate working with legally
restricted or sensitive data (e.g., HIPAA-, FERPA-, or CUI-regulated data), see
the Secure Computing page. Researchers also have the option of participating in
the Condo Cluster Program (CCP). The CCP allows researchers to reserve computing
resources for dedicated use by their projects, giving them the flexibility of
having their own cluster without the responsibility of purchasing hardware or
performing maintenance. For more information on the CCP, see the Condo Cluster
Program page.

## The Penn Letter N-Back Test

The Penn Letter N-Back Test measures working memory (WM), the ability to keep
and refresh goal-related information. Participants attend to a continual series
of letters that flash on the screen (one at a time) and press the spacebar
according to three different rules (called the 0-back, 1-back, and 2-back).
During the 0-back condition, the participant must simply respond to a currently
present target ("X"). During the 1-back condition, he/she must press the
spacebar when the letter on the screen is the same as the previous letter.
During the 2-back condition, he/she must press the spacebar when the letter on
the screen is the same as the letter before the previous letter (i.e. 2 letters
back). In all trials, the inter-stimulus interval (ISI) is 2.5 seconds, and the
stimuli (letters) themselves are presented for 0.5 seconds each. The participant
practices all three principles before testing.
