---
title: "GenomicSEM Analysis g ADHD"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#' Starting from GSM website
##read in the schizophrenia summary statistics using fread from the data.table R package##
library(tidyverse)
library(GenomicSEM)
require(here)
```

# Common Factor GWAS

## Step 1: Munge the summary statistics

```{r, eval=FALSE}
# symlink GWAS summary stats
ln -shfi ~/gwm.flow/data/genotypes/w_hm3.noMHC.snplist ~/gwm.flow/w_hm3.noMHC.snplist 
ln -shfi ~/gwm.flow/data/genotypes/anx.sumstats.gz ~/gwm.flow/anx.sumstats.gz 

## to zcat and see inside zipped files
zcat < anx.sumstats.gz | wc -l
zcat < anx.sumstats.gz | more 

ln -shfi ~/gwm.flow/data/genotypes/anx.sumstats.gz ~/gwm.flow/anx.sumstats.gz 

## to zcat and see inside zipped files
zcat < anx.sumstats.gz | wc -l
zcat < anx.sumstats.gz | more 

```

```{r}
# trying with symlinked files from /data/genotypes
munge(c(
   "daner_cogent_38.filtered.results", # cogent
   "GWAS_CP_all.txt", # from ssgac
   "daner_adhd_meta_filtered_NA_iPSYCH23_PGC11_sigPCs_woSEX_2ell6sd_EUR_Neff_70.meta.gz"
      ), 
   hm3 = "w_hm3.snplist",
   trait.names=c(
      "gcog_1",
      "gcog_2",
      "adhd"
      ), 
   c(
     28445,
     257828,
     55274),
   info.filter = 0.9,
   maf.filter = 0.01) &
```

## Step 2: Run multivariable LDSC

```{r}
traits <- c(
    "cog.sumstats.gz", "adhd.sumstats.gz", "anx.sumstats.gz",
    "mdd.sumstats.gz", "scz.sumstats.gz"
)
sample.prev <- c(NA, 0.37, 0.32, 0.35, 0.39)
population.prev <- c(NA, 0.1, 0.29, 0.16, 0.01)
ld <- "eur_w_ld_chr/"
wld <- "eur_w_ld_chr/"
trait.names <- c("COG", "ADHD", "ANX", "MDD", "SCZ")
stand <- FALSE
# function
LDSCoutput <- GenomicSEM::ldsc(
    traits = traits, sample.prev = sample.prev,
    population.prev = population.prev, ld = ld,
    wld = wld, trait.names = trait.names, stand = stand
)
```


## Step 3: Prepare the summary statistics for GWAS

The sumstats function takes care of a few things prior to running multivariate
GWAS. In order to run a multivariate GWAS, the SNPs, and corresponding SNP
effects, need to be coded across phenotypes so that the same allele is the
reference allele in all cases. In all cases, coefficients and their SEs are then
further transformed such that they are scaled relative to unit-variance scaled
phenotypes. In order to appropriately perform this standardization for different
types of betas (e.g., logistic betas, odds-ratios), the sumstats function takes
a number of arguments outlined below. The sumstats function then combines these
newly scaled betas and SEs into a single, listwise deleted file. Note that the
OLS and LPM transformation require only a total sample size, p-value, and signed
effect column (e.g., beta or Z-statistic), and do not require a SE column.
Transformations for a case/control GWAS require both a SE column and either an
odds ratio or logistic beta. Confused about how to specify the arguments for
this function? See point 3 on the important resources and key information page.
Please also note that the sumstats function requires that the variables be
listed in the same order as they are listed for the ldsc function above. The
sumstats function takes 4 necessary arguments, along with 6 additional arguments
whose default is NULL:

```{r}
# GWAS
files <- c(
    "cogent_g_n38.txt", "adhd_pgc_2019.txt", "anxiety.txt",
    "mdd_2018.txt", "scz_withRS.txt"
)
ref <- "reference.1000G.maf.0.005.txt"
trait.names <- c("COG", "ADHD", "ANX", "MDD", "SCZ")
se.logit <- c(F, F, F, F, F)
OLS <- c(T, F, F, F, F)
linprob <- NULL
prop <- NULL
N <- c(28445, 55274, 17310, 173005, 105318)
info.filter <- 0.6
maf.filter <- 0.01
keep.indel <- FALSE
parallel <- FALSE
cores <- NULL

p_sumstats <- GenomicSEM::sumstats(
    files = files, ref = ref, trait.names = trait.names,
    se.logit = se.logit, OLS = OLS, linprob = linprob, prop = prop,
    N = N, info.filter = info.filter, maf.filter = maf.filter,
    keep.indel = keep.indel, parallel = parallel, cores = cores
)
```

## Step 4: Combine the summary statistics and LDSC output and run the common factor GWAS

Note that as of 4/6/21, we combine the addSNPs and GWAS functions to save
memory. The output from addSNPs can no longer be fed to the GWAS functions. In
order to run multivariate GWAS, the S and V matrices from LDSC and the newly
prepared summary statistics files first need to be merged to create as many S
and V matrices as there are SNPs. The commonfactorGWAS function does this by
expanding the S and V matrices to include the SNP effects. In addition, the
function transforms the GWAS summary statistics (i.e., regression coefficients)
and their SEs into covariances and SEs of covariances by taking the product of
the regression coefficient and SNP variance from the reference file. Finally,
the function will smooth S and V matrices that are non-positive definite when
including the SNP covariances. After constructing the S and V matrices, the
function will run the common factor model.

Note that the commonfactorGWAS function use unit loading identification, such
that the loading of the first indicator is fixed to 1 so that the model is
identified. In order for the model to converge, it will typically be best to
order your files in the ldsc and sumstats function such that the phenotype with
the largest loading on the common factor is listed first. This can be determined
by running commonfactor on just the ldsc matrix to determine what the factor
structure looks like when not including the effect of an individual SNP. In this
case, the phenotype with the largest loading on the common factor is
schizophrenia. The commonfactorGWAS function has the following arguments:

1.  covstruc: The output from LDSC.

2.  SNPs: The output from sumstats.

3.  estimation: Whether the models should be estimated using "DWLS" or "ML"
    estimation. The package default is "DWLS".

4.  cores: How many computer cores to use in parallel processing. The default is
    to use 1 less core than is available in the local environment.

5.  toler: What tolerance level to use for matrix inversions. This is only
    something that needs to be of concern if warning/error messages are produced
    to the effect of "matrix could not be inverted".

6.  SNPSE: An optional argument specifying whether you want to set the SE of the
    SNP to a value higher than the default of .0005. This number is set low to
    reflect that this value is essentially be treating as fixed.

7.  parallel: An optional argument specifying whether you want the function to
    be run in parallel, or to be run serially. As most outputs from sumstats
    will be quite long (i.e., \> 1 million SNPs), running the function in
    parallel is highly recommended in order to save time; the default is for the
    function to run in parallel unless you are using a Windows operating system.

8.  GC: Level of Genomic Control (GC) you want the function to use. The default
    is 'standard' which adjusts the univariate GWAS standard errors by
    multiplying them by the square root of the univariate LDSC intercept.
    Additional options include 'conserv', which adjusts standard errors by
    multiplying with univariate LDSC intercept, and 'none' which does not
    correct the standard errors.

9.  MPI: Whether the function should use multi-node processing (i.e., MPI).
    Please note that this should only be used on a computing cluster on which
    the R package Rmpi is already installed.

```{r}
# commonfactor ------------------------------------------------------------
covstruc <- LDSCoutput
cf <- GenomicSEM::commonfactor(covstruc, estimation = "DWLS")
```

```{r}
# run the multivariate GWAS using parallel processing
covstruc <- LDSCoutput
SNPs <- p_sumstats
estimation <- "DWLS"
cores <- NULL # change for CARC
toler <- FALSE
SNPSE <- FALSE
parallel <- FALSE # change for CARC
GC <- "standard"
MPI <- FALSE
# function
pfactor <- commonfactorGWAS(
    covstruc = covstruc, SNPs = SNPs, estimation = estimation,
    cores = cores, toler = toler, SNPSE = SNPSE, parallel = parallel,
    GC = GC, MPI = MPI
)
```

```{r}
# note that the code written above specifies all arguments for completeness, but as many of these arguments
# are set to the package default it could also be written as below and produce the same result:
pfactor <- commonfactorGWAS(covstruc = LDSCoutput, SNPs = p_sumstats)
```

The figure directly below depicts as a path diagram an example of the model run
for SNP rs4552973: <!-- insert image -->

The full set of summary statistics, along with a ReadMe file, can be found at:
<https://utexas.box.com/s/vkd36n197m8klbaio3yzoxsee6sxo11v> The first 6 columns
are pulled from the file created using sumstats and list the rsID, chromosome,
BP, MAF from the reference file, A1, and A2. The "i" column is simply an ordered
list of the runs. The "lhs", "op", and "rhs" column lists the outcome, operator,
and predictor. In this case "F1 \~ SNP", indicates the common factor regressed
on the SNP. The user should check that this column reads as such to confirm the
appropriate results are being pulled. The "est" column lists the effect of the
SNP on the common factor. The "se_c" column lists the sandwich corrected
standard error associated with this effect. The "Z_Estimate" column lists the
ratio of est/se_c. The "Pval_Estimate" column lists the corresponding p-value of
Z_Estimate.

The Q column lists the heterogeneity estimate for that SNP. This indexes the
extent to which the SNP effect is not mediated by the common factor, with larger
values indicating greater heterogeneity. This is a chi-square distributed test
statistic, with degrees of freedom for a common factor model equal to the number
of indicators - 1. The "Q_df" column lists the degrees of freedom for Q, and the
"Q_pval" column lists the p-value associated with the Q statistic. The fail
column indicates whether a particular run did not converge. The warning column
will list whether a particular run raised any warnings in lavaan. This may
include specific runs where the residual variance of an indicator is negative.

Calculating sample size for factors Once the summary statistics for a common
factor are produced, the user may wish to input those summary statistics into an
outside program that requires a sample size (e.g., LDSC, LDpred). A description
of one method for calculating effective sample size can be found in the online
supplement of bioRxiv pre-print from Mallard et al. (2019):
<https://www.biorxiv.org/content/10.1101/603134v1.abstract>. The equation
described in the supplement can be written as below to produce the effective
sample size for the p-factor. Note that the recommendation is also made to
restrict the summary statistics to MAF limits of approximately 10% and 40% in
order to produce more stable estimates.

```{r}
# restrict to MAF of 40% and 10%
pfactor2 <- subset(pfactor, pfactor$MAF <= 0.4 & pfactor$MAF >= 0.1)

# calculate effective N
Effective_N <- (mean(
    ((pfactor2$Z_Estimate/pfactor2$est)^2)/(2 * pfactor2$MAF * (1 -
        pfactor2$MAF))
)) &
```

We note also that for LDSC, the sample size provided does not effect the
univariate intercept or the intercept SE. Although the heritability estimate and
SE are contingent on user provided sample size, the ratio of the heritability
estimate over SE is retained regardless of the sample size provided and,
therefore, the heritability z-statistic can still be calculated.

The commonfactorGWAS and commonfactorGWASpar functions use unit loading
identification and genetic covariance matrices (and corresponding sampling
covariance matrices) as input. That is, the SNP effects shown in the figure and
table above are unstandardized with respect to both the outcome (the common
factor) and the predictor (the SNP). As the common factor is a dependent
variable, its variance is not directly represented, but rather implied. Thus,
unit variance identification would require a model constraint. The latest
version includes a user defined GWAS function in which the the user has
substantial flexibility regarding what covariance structure and parameterization
(such as specifying a model constraint) is specified. Output for a fully
standardized version of a common factor GWAS is shown as an example below for
rs4552973:

A note on serial versus parallel processing. Serial and parallel processing will
return the same output, but are computationally implemented in a serial and
parallel fashion, respectively. In our experience using high performance
computing systems there were often memory issues when using parallel processing
unless you have access to large memory nodes. If memory issues are a problem
then serial processing may be preferred. However, on a personal laptop with
multiple cores, the computing time will be reduced greatly when using parallel
processing. On a personal laptop with 8 cores using parallel processing for a
common factor with five indicators and 1 million SNPs, the run time was
approximately 8 hours. These functions are running as many models as there are
SNPs.




# 3. Models without Individual SNP effects

## Step 1: Munge the summary statistics

```{r}
##read in the schizophrenia summary statistics using fread from the data.table R package##
require(data.table)
scz<-fread("clozuk_pgc2.meta.sumstats.txt",data.table=FALSE) 
##pull out the rsIDs##
scz$SNP<-sapply(strsplit(scz$SNP, ":"), `[`, 1)
##output the result in a .txt file called scz_withRS.txt##
write.table(scz, file = "scz_withRS.txt", sep = "\t", quote = FALSE, row.names = FALSE, col.names=TRUE)
##munge the summary statistics files##
##note that we include the argument names (e.g., files, hm3) for
##completeness in the code below, but that this is not necessary to run the code.
munge(files=c("scz_withRS.txt", "pgc.bip.full.2012-04.txt", "MDD2018_ex23andMe.txt", "SORTED_PTSD_EA9_ALL_study_specific_PCs1.txt", "anxiety.meta.full.cc.tbl"), hm3 = "w_hm3.snplist",trait.names=c("scz", "bip", "mdd", "ptsd", "anx"),N=c(105318, 16731, 173005, 9537, 17310), info.filter = 0.9, maf.filter = 0.01)
```

```{r}
#GWM
scz<-fread("clozuk_pgc2.meta.sumstats.txt",data.table=FALSE) 

```

## Step 2: Run multivariable LDSC

```{r}
traits <- c("scz.sumstats.gz", "bip.sumstats.gz", "mdd.sumstats.gz", "ptsd.sumstats.gz", "anx.sumstats.gz")
sample.prev <- c(.39,.45,.35,.25,.32)
population.prev <- c(.01,.01,.16,.08,.29)
ld <- "eur_w_ld_chr/"
wld <- "eur_w_ld_chr/"
trait.names<-c("SCZ","BIP","MDD","PTSD","ANX")
LDSCoutput <- ldsc(traits, sample.prev, population.prev, ld, wld, trait.names)

##optional command to save the ldsc output in case you want to use it in a later R session. 
save(LDSCoutput, file="Pfactor.RData")

k<-nrow(LDSCoutput$S)
SE<-matrix(0, k, k)
SE[lower.tri(SE,diag=TRUE)] <-sqrt(diag(LDSCoutput$V))


```

## Common Factor Model

```{r}
#To run using DWLS estimation#
CommonFactor_DWLS<- commonfactor(covstruc = LDSCoutput, estimation="DWLS")

#print CommonFactor_DWLs output#
CommonFactor_DWLS

```

```{r}
commonfactor.model<-'F1=~ NA*SCZ + BIP + MDD + PTSD + ANX
F1~~1*F1
```

## User-Specified Models

Genomic SEM is not limited to common factor models, and can be used to fit any
form of structural model, enabling the ability to test a myriad of models and
theoretical accounts. We outline some examples of how to specify your own models
here, starting with the GWIS example from the paper.

### Genetic multiple regression (similar to GWIS)

```{r}
#munge the files
munge(c("pgc.cross.SCZ17.2013-05.txt", "pgc.cross.BIP11.2013-05.txt", "SSGAC_EduYears_Rietveld2013_publicrelease.txt"), hm3 = "w_hm3.snplist",trait.names=c("sczcross", "bipcross", "EA"), c(26552, 28675, 126559), info.filter = 0.9, maf.filter = 0.01)

#run multivariable LDSC to create the S and V matrices
ld <- "eur_w_ld_chr/"
wld <- "eur_w_ld_chr/"
traits <- c("sczcross.sumstats.gz","bipcross.sumstats.gz", "EA.sumstats.gz")
sample.prev <- c(.27,.32,NA)
population.prev <- c(.01,.01,NA)
trait.names<-c("SCZ", "BIP", "EA")

GWIS<-ldsc(traits, sample.prev, population.prev, ld, wld, trait.names)

```

```{r}
#specify a model in which Educational Achievement is regressed on Bipolar Disorder and Schizophrenia, and the correlation is modeled #between Bipolar and Schizophrenia
GWISmodel <- 'EA ~ SCZ + BIP
SCZ~~BIP'

#run the model using the usermodel function
GWISoutput<-usermodel(covstruc = GWIS, estimation = "DWLS", model = GWISmodel, CFIcalc = TRUE, std.lv = FALSE, imp_cov = FALSE)

#print the output
GWISoutput
```

## Exploratory Factor Analysis

```{r}
#munge the summary statistics
munge(c("SNP_gwas_mc_merge_nogc.tbl.uniq", "GIANT_2015_WHR_COMBINED_EUR.txt", "EGG_Obesity_Meta_Analysis_1.txt", "GIANT_2015_HIP_COMBINED_EUR.txt", "GIANT_2015_WC_COMBINED_EUR.txt", "GIANT_HEIGHT_Wood_et_al_2014_publicrelease_HapMapCeuFreq.txt", "EGG_HC_DISCOVERY.v2.txt", "EGG-GWAS-BL.txt", "EGG_BW2_DISCOVERY.txt"), hm3 = "w_hm3.snplist",trait.names=c("BMI2015", "waisthip", "childobese", "hip", "waist", "height", "headcirc", "birthlength", "birthweight"), c(NA, NA, 13848, NA, NA, NA, NA, NA, 26836), info.filter = 0.9, maf.filter = 0.01)

#run multivariable LDSC to create the S and V matrices
ld <- "eur_w_ld_chr/"
wld <- "eur_w_ld_chr/"
traits<-c("BMI2015.sumstats.gz", "waisthip.sumstats.gz", "childobese.sumstats.gz", "waist.sumstats.gz","hip.sumstats.gz", "height.sumstats.gz", "headcirc.sumstats.gz", "birthlength.sumstats.gz", "birthweight.sumstats.gz")
sample.prev <- c(NA,NA,NA,NA,NA,NA,NA,NA,NA)
population.prev <- c(NA,NA,NA,NA,NA,NA,NA,NA,NA)
trait.names<-c("BMI","WHR","CO","Waist", "Hip", "Height", "IHC", "BL", "BW")
anthro<-ldsc(traits, sample.prev, population.prev, ld, wld, trait.names)

#smooth the S matrix for EFA using the nearPD function in the Matrix package. 
require(Matrix)
Ssmooth<-as.matrix((nearPD(anthro$S, corr = FALSE))$mat)

#run EFA with promax rotation and 2 factors using the factanal function in the stats package
require(stats)
EFA<-factanal(covmat = Ssmooth, factors = 2, rotation = "promax")

#print the loadings
EFA$loadings

```

## Confirmatory factor analysis

```{r}
#Specify the Genomic confirmatory factor model
CFAofEFA <- 'F1 =~ NA*BMI + WHR + CO + Waist + Hip
             F2 =~ NA*Hip + Height + IHC + BL + BW
F1~~F2
Waist ~~ a*Waist
a > .001'

#run the model
Anthro<-usermodel(anthro, estimation = "DWLS", model = CFAofEFA, CFIcalc = TRUE, std.lv = TRUE, imp_cov = FALSE)

#print the results
Anthro
```
