---
title: "GWAS of the RDoC Cognitive Systems Domain: Modeling the Latent Genetic Architecture of Working Memory"
subtitle: "B.2 Accomplishments"
author: "Joey Trampush PhD"
date: "`r Sys.Date()`"
output: tint::tintPdf
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tint)
# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tint'))
options(htmltools.dir.version = FALSE)
```

# Major activities

# Specific objectives

# Significant results

- including major findings, developments, or conclusions (both positive and negative)

# Key outcomes or other achievements


Include a discussion of stated goals not met. As the project progresses, the
emphasis in reporting in this section should shift from reporting activities to
reporting accomplishments. Goals are equivalent to specific aims. In the response, emphasize the significance of the findings to the scientific field. Include the approaches taken to ensure robust and unbiased results. For most NIH awards the response should not exceed 2 pages.


## High-Performance Computing

Computational research often requires resources that exceed those of a personal
laptop or desktop computer. High-performance computing aggregates the resources
from individual computers (known as nodes) into a cluster that works together to perform advanced, specialized computing jobs. The Center for Advanced Research
Computing launched its new high-performance computing cluster, Discovery, in
August 2020. The new cluster marks a significant upgrade to the CARC's
cyberinfrastructure, and the first step in a major, user-focused overhaul of the
program. The Discovery cluster consists of 2 shared login nodes and a total of
around 11,000 CPU cores in around 500 compute nodes. The typical compute node
has dual 8 to 16 core processors and resides on a 56 gigabit FDR InfiniBand
backbone. Nearly 100 GB of storage space is provided for each user to store
important code and configuration files in their home directories, and the
BeeGFS/ZFS parallel project file system has a capacity of 8.4 PB of usable
space, with a default storage quota of 10 TB per Principal Investigator across
their projects. A rapidly growing fleet of state-of-the art multicore compute
nodes is available for data-intensive research jobs, backed by two 800 TB,
high-throughput scratch file systems. A low-latency, high-bandwidth InfiniBand
network fabric facilitates intense research workloads on large data sets.
Discovery includes an array of scientific software packages, both licensed and
open source, for engineering, molecular simulation, and computational chemistry.
Researchers can also install software packages or develop their own code within
their project's allotted storage. Discovery is free to use for all USC faculty,
research staff, and graduate students (with the approval of their faculty
advisor). For information on how to set up your account and start using the
cluster, see the Accounts page. If you anticipate working with legally
restricted or sensitive data (e.g., HIPAA-, FERPA-, or CUI-regulated data), see
the Secure Computing page. Researchers also have the option of participating in
the Condo Cluster Program (CCP). The CCP allows researchers to reserve computing
resources for dedicated use by their projects, giving them the flexibility of
having their own cluster without the responsibility of purchasing hardware or
performing maintenance. For more information on the CCP, see the Condo Cluster
Program page.

## The Penn Letter N-Back Test

The Penn Letter N-Back Test measures working memory (WM), the ability to keep
and refresh goal-related information. Participants attend to a continual series
of letters that flash on the screen (one at a time) and press the spacebar
according to three different rules (called the 0-back, 1-back, and 2-back).
During the 0-back condition, the participant must simply respond to a currently
present target ("X"). During the 1-back condition, he/she must press the
spacebar when the letter on the screen is the same as the previous letter.
During the 2-back condition, he/she must press the spacebar when the letter on
the screen is the same as the letter before the previous letter (i.e. 2 letters
back). In all trials, the inter-stimulus interval (ISI) is 2.5 seconds, and the
stimuli (letters) themselves are presented for 0.5 seconds each. The participant
practices all three principles before testing.


## Margin Figures

Images and graphics play an integral role in Tufte's work. To place figures in
the margin you can use the **knitr** chunk option `fig.margin = TRUE`. For
example:

```{r fig-margin, fig.margin = TRUE, fig.cap = "MPG vs horsepower, colored by transmission.", fig.width=3.5, fig.height=3.5, cache=TRUE}
library(ggplot2)
mtcars2 <- mtcars
mtcars2$am <- factor(
  mtcars$am, labels = c('automatic', 'manual')
)
ggplot(mtcars2, aes(hp, mpg, color = am)) +
  geom_point() + geom_smooth() +
  theme(legend.position = 'bottom')
```

Note the use of the `fig.cap` chunk option to provide a figure caption. You can
adjust the proportions of figures using the `fig.width` and `fig.height` chunk
options. These are specified in inches, and will be automatically scaled down to
fit within the handout margin.

## Arbitrary Margin Content

In fact, you can include anything in the margin using the **knitr** engine named
`marginfigure`. Unlike R code chunks ```` ```{r} ````, you write a chunk
starting with ```` ```{marginfigure} ```` instead, then put the content in the
chunk. See an example on the right about the first fundamental theorem of
calculus.

```{marginfigure}
We know from _the first fundamental theorem of calculus_ that for $x$ in $[a, b]$:
$$\frac{d}{dx}\left( \int_{a}^{x} f(u)\,du\right)=f(x).$$
```

For the sake of portability between LaTeX and HTML, you should keep the margin
content as simple as possible (syntax-wise) in the `marginefigure` blocks. You
may use simple Markdown syntax like `**bold**` and `_italic_` text, but please
refrain from using footnotes, citations, or block-level elements (e.g.
blockquotes and lists) there.

## Full Width Figures

You can arrange for figures to span across the entire page by using the chunk
option `fig.fullwidth = TRUE`.

```{r fig-fullwidth, fig.width = 10, fig.height = 2, fig.fullwidth = TRUE, fig.cap = "A full width figure.", warning=FALSE, cache=TRUE}
ggplot(diamonds, aes(carat, price)) + geom_smooth() +
  facet_grid(~ cut)
```

Other chunk options related to figures can still be used, such as `fig.width`,
`fig.cap`, `out.width`, and so on. For full width figures, usually `fig.width`
is large and `fig.height` is small. In the above example, the plot size is
$10 \times 2$.

## Main Column Figures

Besides margin and full width figures, you can of course also include figures
constrained to the main column. This is the default type of figures in the
LaTeX/HTML output.

```{r fig-main, fig.cap = "A figure in the main column.", cache=TRUE}
ggplot(diamonds, aes(cut, price)) + geom_boxplot()
```

# Sidenotes

One of the most prominent and distinctive features of this style is the
extensive use of sidenotes. There is a wide margin to provide ample room for
sidenotes and small figures. Any use of a footnote will automatically be
converted to a sidenote. [^1]

[^1]: This is a sidenote that was entered using a footnote.

If you'd like to place ancillary information in the margin without the sidenote
mark (the superscript number), you can use the `margin_note()` function from
**tufte** in an inline R expression.
`r margin_note("This is a margin note.  Notice that there is no number preceding the note.")`
This function does not process the text with Pandoc, so Markdown syntax will not
work here. If you need to write anything in Markdown syntax, please use the
`marginfigure` block described previously.

# Tables

You can use the `kable()` function from the **knitr** package to format tables
that integrate well with the rest of the Tufte handout style. The table captions
are placed in the margin like figures in the HTML output.

```{r}
knitr::kable(
  mtcars[1:6, 1:6], caption = 'A subset of mtcars.'
)
```


