---
title: "GWAS of the RDoC Cognitive Systems Domain: Modeling the Latent Genetic Architecture of Working Memory"
subtitle: "B.2 Accomplishments"
author: "Joey Trampush PhD"
date: "`r Sys.Date()`"
output: tint::tintPdf
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tint)
# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tint'))
options(htmltools.dir.version = FALSE)
```

# Major activities

# Specific objectives

# Significant results

-   including major findings, developments, or conclusions (both positive and
    negative)

# Key outcomes or other achievements

Include a discussion of stated goals not met. As the project progresses, the
emphasis in reporting in this section should shift from reporting activities to
reporting accomplishments. Goals are equivalent to specific aims. In the
response, emphasize the significance of the findings to the scientific field.
Include the approaches taken to ensure robust and unbiased results. For most NIH
awards the response should not exceed 2 pages.

## Margin Figures

Images and graphics play an integral role in Tufte's work. To place figures in
the margin you can use the **knitr** chunk option `fig.margin = TRUE`. For
example:

```{r fig-margin, fig.margin = TRUE, fig.cap = "MPG vs horsepower, colored by transmission.", fig.width=3.5, fig.height=3.5, cache=TRUE}
library(ggplot2)
mtcars2 <- mtcars
mtcars2$am <- factor(
  mtcars$am, labels = c('automatic', 'manual')
)
ggplot(mtcars2, aes(hp, mpg, color = am)) +
  geom_point() + geom_smooth() +
  theme(legend.position = 'bottom')
```

Note the use of the `fig.cap` chunk option to provide a figure caption. You can
adjust the proportions of figures using the `fig.width` and `fig.height` chunk
options. These are specified in inches, and will be automatically scaled down to
fit within the handout margin.

## Arbitrary Margin Content

In fact, you can include anything in the margin using the **knitr** engine named
`marginfigure`. Unlike R code chunks ```` ```{r} ````, you write a chunk
starting with ```` ```{marginfigure} ```` instead, then put the content in the
chunk. See an example on the right about the first fundamental theorem of
calculus.

```{marginfigure}
We know from _the first fundamental theorem of calculus_ that for $x$ in $[a, b]$:
$$\frac{d}{dx}\left( \int_{a}^{x} f(u)\,du\right)=f(x).$$
```

For the sake of portability between LaTeX and HTML, you should keep the margin
content as simple as possible (syntax-wise) in the `marginefigure` blocks. You
may use simple Markdown syntax like `**bold**` and `_italic_` text, but please
refrain from using footnotes, citations, or block-level elements (e.g.
blockquotes and lists) there.

## Full Width Figures

You can arrange for figures to span across the entire page by using the chunk
option `fig.fullwidth = TRUE`.

```{r fig-fullwidth, fig.width = 10, fig.height = 2, fig.fullwidth = TRUE, fig.cap = "A full width figure.", warning=FALSE, cache=TRUE}
ggplot(diamonds, aes(carat, price)) + geom_smooth() +
  facet_grid(~ cut)
```

Other chunk options related to figures can still be used, such as `fig.width`,
`fig.cap`, `out.width`, and so on. For full width figures, usually `fig.width`
is large and `fig.height` is small. In the above example, the plot size is
$10 \times 2$.

## Main Column Figures

Besides margin and full width figures, you can of course also include figures
constrained to the main column. This is the default type of figures in the
LaTeX/HTML output.

```{r fig-main, fig.cap = "A figure in the main column.", cache=TRUE}
ggplot(diamonds, aes(cut, price)) + geom_boxplot()
```

# Sidenotes

One of the most prominent and distinctive features of this style is the
extensive use of sidenotes. There is a wide margin to provide ample room for
sidenotes and small figures. Any use of a footnote will automatically be
converted to a sidenote. [^1]

[^1]: This is a sidenote that was entered using a footnote.

If you'd like to place ancillary information in the margin without the sidenote
mark (the superscript number), you can use the `margin_note()` function from
**tufte** in an inline R expression.
`r margin_note("This is a margin note.  Notice that there is no number preceding the note.")`
This function does not process the text with Pandoc, so Markdown syntax will not
work here. If you need to write anything in Markdown syntax, please use the
`marginfigure` block described previously.

# Tables

You can use the `kable()` function from the **knitr** package to format tables
that integrate well with the rest of the Tufte handout style. The table captions
are placed in the margin like figures in the HTML output.

```{r}
knitr::kable(
  mtcars[1:6, 1:6], caption = 'A subset of mtcars.'
)
```
