---
output:
  bookdown::pdf_document2:
    toc: no
    template: template_nih.tex
    keep_tex: yes
    keep_md: yes
    number_sections: no
    latex_engine: xelatex
    citation_package: natbib
    fig_crop: yes
    fig_caption: yes
site: bookdown::bookdown_site
geometry: margin=0.5in
mainfont: Fira Sans
fontsize: 11pt
indent: false
bibliography: gwm.bib
link-citations: yes
colorlinks: yes
graphics: yes
fig-caption: yes
self-contained: yes
highlight: yes
linestretch: 1.24
---

```{r, setup, include = FALSE, results='hide', warning=FALSE}
library(knitr)
knitr::knit_hooks$set(crop = knitr::hook_pdfcrop)
knitr::knit_hooks$set(optipng = knitr::hook_optipng)

# Fix subfigs for rmarkdown -> latex
if (identical(knitr:::pandoc_to(), 'latex'))
  knitr::knit_hooks$set(plot = knitr::hook_plot_tex)

knitr::opts_chunk$set(
  echo = FALSE,
  dev = c("png", "pdf"),
  dev.args = list(pointsize = 8),
  crop = TRUE,
  par = TRUE,
  fig.width = 6.5,
  fig.height = 5.5,
  dpi = 300,
  #fig.pos = '!H',
  root.dir = normalizePath('./'),
  cache = FALSE,
  cache.path = './cache/report-',
  # Code block options
  eval = TRUE,
  include = TRUE,
  message = FALSE,
  warning = FALSE,
  results = 'asis',
  error = FALSE,
  tidy = TRUE
)

options(digits = 3, warnPartialMatchArgs = FALSE)
```

## B.2 What was accomplished under these goals?

Year 1 of the project was challenged by the COVID-19 pandemic and slowed our
launch to a degree. We are catching up now, but we had restricted access to our
labs at USC for almost the entire year and limited access to
people/colleagues/students. We weathered it as best we could and thankfully are
back on track, and are looking forward to completing our study aims in Year 2.

### Reproducible research workflow

The past few years have yielded much discussion and controversy in biomedical
research circles surrounding reproducibility and replication of findings
[@Krieger2019]. This is a well known problemm in psychiatric genetics. Thus,
prior to engaging in the core data analyses for the study, I deliberately carved
out time in my schedule for several months while working remotely to learn more
about this issue and made a concerted effort to develop a consistent analytic
workflow. I focused mostly on the open-source tools provided by the R
statistical software package and specifcally learned how to work wiht and use
the R-based "Tidyverse" ecosystem [ref], which is a set of tools that supports
an efficient project management workflow for statisticians and data scientists
who perform large-scale reproducible research within team science environments.
The R Tidyverse is one of the leading software systems in the growing ecosystem
of reproducibility-facilitating software for data science. I have designed my
labs reinvented analtyical strategy for this project as well as all future
projects to work in concert with RStudio, an interactive R integrated
development environment [9], and R Markdown (.Rmd), a file structure
accommodating text, code and code output [10]. Fully integrated with RStudio, R
Markdown allows us to interleave statistical programming code with formatted
annotations, enabling its users to write highly technical documents, perform
cutting-edge statistical analyses with interaractive data visualzations, apply
version control via Git and GitHub, publish our methods in publicaly available
open-scienece frameworks and repositories, share our code, and generate
easy-to-use websites were all of our methods can be easily shared and reviewed
by any qualified and/or interested researcerh. We have a website for the current
project at [gwm.flow](https://brainworkup.github.io/gwm.flow).

For example, I generated these visualtions of working memory.

```{r}
## Create Ridgeline plot using `ggridges`
# Load required R packages:
library(here)
library(tidyverse)
library(data.table)
library(purrr)
library(ggridges) # ridgeline plots
library(ggthemes) # wsj, tufte, 538, hc
library(gganimate) # animate and create gifs
library(scales)
library(tweenr)
library(animation)
library(htmlwidgets)
library(widgetframe)
library(crosstalk)
library(manipulateWidget)
```

```{r eval=TRUE}
# NIH EXAMINER working memory data - fill sample
#wm <- data.table::fread("data/wm.csv")
wm <- data.table::fread("data/wm_10-80.csv")

# wm phenos
pheno <- data.frame(wm %>% select(
  age_10y,
  wm,
  dots,
  nback2,
  nback1)
  )

dftmp <- as.data.frame(scale(pheno))
names(dftmp) <- paste0("z", names(dftmp))
pheno <-  bind_cols(pheno, dftmp)

pheno <- data.frame(pheno %>% select(
  age_10y,
  zwm,
  zdots,
  znback2,
  znback1)
  )

# Define function using .data[[]] pronoun

plot_for_loop <- function(df, x_var, y_var) {

  # plot code
  plots <- ggplot(df, na.rm = TRUE, aes(x = .data[[x_var]], y = as.factor(.data[[y_var]]), fill = stat(x))) +
    geom_density_ridges_gradient(scale = 3, size = 0.3, rel_min_height = 0.01, gradient_lwd = 0.5) +
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_discrete(expand = expansion(mult = c(0, 0))) +
    scale_fill_viridis_c(name = "score\nrange:", option = "A", direction = -1) +
    labs(
      title = x_var,
      subtitle = "",
      x = "z-score",
      y = "age group") +
    #theme_fivethirtyeight() +
    #theme_wsj() +
    theme(axis.title = element_text(size = 10))
}

# Loop through every column
plot_list <- colnames(pheno)[-1] %>%
  map( ~ plot_for_loop(pheno, .x, colnames(pheno)[1]))

# save each file to the corresponding location on disk:
paths <- stringr::str_c(names(pheno)[-1], ".pdf")
pwalk(list(paths, plot_list), ggsave, path = getwd(), device = "pdf", scale = 1)

# indiv plots
p1 <- plot_list[[1]] +
   ggtitle('WM factor')
p2 <- plot_list[[2]] +
  ggtitle('dot counting')
p3 <- plot_list[[3]] +
  ggtitle('nBack1')
p4 <- plot_list[[4]] +
  ggtitle('nBack2')
```

```{r}
defOut <- knitr::knit_hooks$get("plot")  # save the default plot hook
knitr::knit_hooks$set(plot = function(x, options) {  # set new plot hook ...
  x <- defOut(x, options)  # first apply the default hook
  if(!is.null(options$wrapfigure)) {  # then, if option wrapfigure is given ...
    # create the new opening string for the wrapfigure environment ...
    wf <- sprintf("\\begin{wrapfigure}{%s}{%g\\textwidth}", options$wrapfigure[[1]], options$wrapfigure[[2]])
    x  <- gsub("\\begin{figure}", wf, x, fixed = T)  # and replace the default one with it.
    x  <- gsub("{figure}", "{wrapfigure}", x, fixed = T)  # also replace the environment ending
  }
  return(x)
})
```

```{r wm-plot, fig.cap = 'Complex working memory skills (e.g., performance on the dot-counting task and the 2-back trial of the N-back) rapidly improve during childhood/adolescence, peak between ages 20-40, and then decline back to early childhood levels in the later decades of life.', fig.width = 8, fig.height = 7, out.width = ".55\\textwidth", fig.align = "left", wrapfigure = list("L", .55)}

library(patchwork)
p1 /
(p2 | p4 | p3)
```

Related, For the past two months, I have been training on the use of software
containers and singularity. Singularity is an open-source application for
creating and running software containers, designed primarily for
high-performance computing on shared Linux-based computing clusters like CARC
systems.Software containers are isolated, secure, stable, portable, and
reproducible software environments rapidly being embraced by the research
community in an effort to improve research reproducibility. Containers (e.g.,
Docker, Singularity) packages main applications and all dependencies needed for
an analysis, provide OS-level virtualization, and provide a custom user space
that's portable and usable by others almost seamlessly. This ensures the same
software stack is used among a research group, allows the same software stack
across Linux systems (e.g., any HPC center), and one can run the same workflows
across Linux systems by embedding runscripts in container images. A container
image is a single executable file that defines the software environment for a
container and also runs the container. A single container image can be used to
run multiple instances of the same container simultaneously for different jobs.

Singularity containers provide a custom user space and enable isolated, secure,
stable, portable, and reproducible software environments on Linux systems. A
Singularity container bundles a primary application and all of its dependencies
into a single image file, which can also include data, scripts, and other files
if desired. In addition, Singularity containers have direct access to the Linux
kernel on the host system (e.g., Discovery or Endeavour compute nodes), so there
is no substantial performance penalty when using a container compared to using
natively installed software on the host system.

### High-performance computing cluster

In August of 2020, the USC Center for Advanced Research Computing (CARC)
launched a new high-performance computing cluster, Discovery, which is what we
use to compute our large-scale genomic analyses. The new cluster now consists of
around 11,000 CPU cores in around 500 compute nodes; the typical compute node
has dual 8 to 16 core processors and resides on a 56 gigabit FDR InfiniBand
backbone; and nearly 100 GB of storage space is provided for each user, with a
default storage quota of 10 TB per PI across our projects. However, as expected,
the new cluster has had its share of ups and downs in terms of operating
smoothly and wihtout interruption (e.g., it has been offline for maintenance for
several weeks at a time here and there this year, just to work out some of the
unforseeable bugs). There was also a bit of a learning curve to orient to the
new user interfaces and job submission protocols. To support advanced computing
at the university, USC recently upgraded the network capacity of individual
research buildings to 100 Gbps. The CARC has also deployed multiple 100 Gbps
links between the ITS data center and the rest of the campus, creating an
aggregate bandwidth of up to 400 Gbps. CARC will ultimately be an invaluable
resource for computational researchers at USC, but the newness of it just added
some additional wait time for current project.

### GWAS imputation

Finally, the Trans-Omics for Precision Medicine (TOPMed) programme seeks to
elucidate the genetic architecture and biology of heart, lung, blood and sleep
disorders, with the ultimate goal of improving diagnosis, treatment and
prevention of these diseases. The initial phases of the programme focused on
whole-genome sequencing of individuals with rich phenotypic data and diverse
backgrounds. Here we describe the TOPMed goals and design as well as the
available resources and early insights obtained from the sequence data. The
resources include a variant browser, a genotype imputation server, and genomic
and phenotypic data that are available through dbGaP (Database of Genotypes and
Phenotypes)1. In the first 53,831 TOPMed samples, we detected more than 400
million single-nucleotide and insertion or deletion variants after alignment
with the reference genome. Additional previously undescribed variants were
detected through assembly of unmapped reads and customized analysis in highly
variable loci. Among the more than 400 million detected variants, 97% have
frequencies of less than 1% and 46% are singletons that are present in only one
individual (53% among unrelated individuals). These rare variants provide
insights into mutational processes and recent human evolutionary history. The
extensive catalogue of genetic variation in TOPMed studies provides unique
opportunities for exploring the contributions of rare and noncoding sequence
variants to phenotypic variation. Furthermore, combining TOPMed haplotypes with
modern imputation methods improves the power and reach of genome-wide
association studies to include variants down to a frequency of approximately
0.01%.

|                     | **1000G Phase 3** |  **HRC**   | **TOPMed**  |
|:--------------------|:-----------------:|:----------:|:-----------:|
| **Haplotypes**      |       5,008       |   64,940   |   194,512   |
| **Samples**         |       2,504       |   32,470   |   97,256    |
| **Sites (chr1-22)** |    49,143,605     | 39,635,008 | 308,107,085 |

## B.3 Competitive Revisions/Administrative Supplements

For this reporting period, is there one or more Revision/Supplement associated
with this award for which reporting is required?

No

## B.4 What opportunities for training and professional development has the project provided?

Many many. Reproducible research. R. Rmarkdown. GitHub. Git.

Support and Training User support for advanced computing at USC is primarily
available through the CARC, which consists of a systems team, cloud/DevOps team,
and research support team. The staff has expertise in storage system
administration, software development, science and engineering applications,
high-performance and parallel programming, and cloud computing solution
development. The CARC also works with researchers who do not require HPC-level
performance but need help developing or optimizing their code, improving their
computing productivity by learning better programming languages, scripting, and
so forth. The CARC offers regular onboarding sessions, office hours,
introductory and advanced training workshops, and in-person and remote
consultations. It also hosts vendor and national supercomputing center training
at USC. As XSEDE Campus Champions, CARC staff can connect researchers with
high-end computing needs to national supercomputing centers to scale their
workflows and reach other researchers nationwide.

## B.5 How have the results been disseminated to communities of interest?

Neuropsychopharmacology

## B.6 What do you plan to do during the next reporting period to accomplish the goals?

# F SOMETHING

## F.2 Actual or anticipated challenges or delays and actions or plans to resolve them

We are not delayed in meeting any of the aims of the project; however, we have
been impacted by the COVID-19 pandemic. The college students we planned to
include in a beta-test of the intervention in the spring of 2020 were all
off-campus and involved in distance learning. Thus, a beta-test of the pregaming
intervention, which is focused on social drinking, was not possible during the
spring with no students on campus. Students are returning to campus in August
2020, and we will run the beta-test then, followed soon by the launch of the
RCT. We expect drinking contexts to be different in the fall as students return
to campus (e.g., fewer mass gatherings, unclear if football games will be played
with spectators); however, given that some studies are indicating alcohol use
may be increasing for many people during the pandemic, we believe fall 2020 will
be a good time to launch. For students, this will represent the first time they
are getting together with their college friends in months and we expect drinking
to be heavy, and therefore, appropriate to target, in the intervention test.
When reporting findings, we will place the intervention and its effects in
context so that consumers of our study will understand that the intervention was
studied during the COVID-19 pandemic.
