---
output: 
   pdf_document: 
      template: template_nih.tex
      keep_tex: true
      keep_md: true
      number_sections: false
      latex_engine: xelatex
      citation_package: natbib
geometry: margin=.5in
mainfont: Fira Sans #Alegreya Sans # DejaVu Sans Mono for Powerline Book
fontsize: 11pt
indent: true
subparagraph: true
bibliography: references.bib
---

<!-- a clever hack for wrapfig from https://stackoverflow.com/questions/54100806/r-markdown-how-do-i-make-text-float-around-figures -->

```{r, include = F}
defOut <- knitr::knit_hooks$get("plot")  # save the default plot hook 
knitr::knit_hooks$set(plot = function(x, options) {  # set new plot hook ...
  x <- defOut(x, options)  # first apply the default hook
  if(!is.null(options$wrapfigure)) {  # then, if option wrapfigure is given ...
    # create the new opening string for the wrapfigure environment ...
    wf <- sprintf("\\begin{wrapfigure}{%s}{%g\\textwidth}", options$wrapfigure[[1]], options$wrapfigure[[2]])
    x  <- gsub("\\begin{figure}", wf, x, fixed = T)  # and replace the default one with it.
    x  <- gsub("{figure}", "{wrapfigure}", x, fixed = T)  # also replace the environment ending
  }
  return(x)
})
```

#### **NIMH Grant**: 1 R03 MH123787-01 {.unnumbered}

#### **Title**: GWAS of the RDoC Cognitive Systems Domain: Modeling the Latent Genetic Architecture of Working Memory {.unnumbered}

#### **PI:** Joey Trampush {.unnumbered}

#### **Project Period**: 06/15/2020 -- 06/14/2022 {.unnumbered}

This grant was funded in response to NIMH PAR-17-158, "Secondary Data Analyses
to Explore NIMH Research Domain Criteria RDoC (R03)." This report specifies
activities from June 15, 2020 through June 14, 2021.

# B. Accomplishments

## B.1 What are the major goals of the project? {.unnumbered}

**Working Memory**, the most extensively studied of the executive functions, is
the cognitive system responsible for the transient holding and processing of
information in conscious awareness just long enough to use it for problem
solving. Limited working memory faculties are a fundamental aspect of the
cognitive impairments prevalent in ADHD, schizophrenia, and many other CNS
disorders. More than half of the variability in working memory capacity from
person to person can be sourced back to inherited genetic factors. Because it is
so highly heritable and diagnostically unbounded, the NIMH Research Domain
Criteria (RDoC) initiative has prioritized working memory as a core paradigmatic
construct to accelerate gene discovery research. Yet, to date, very few
genome-wide significant findings have been reported for working memory, and none
have been replicated.

As the discovery of loci reliably associated with working memory is key to
understanding neuropathophysiology, our goal is to fast-track the identification
and localization of molecular genetic signatures of working memory, and then to
uncover the underlying causal molecular and biological mechanisms linking
working memory deficits to genetically correlated disorders. GWAS (genome-wide
association study) is a powerful approach to identify molecular signatures of
complex traits if adequate sample sizes and proper statistical procedures are
employed. We will utilize existing genomic and phenotypic data from the
Cognitive Genomics Consortium (COGENT) to test novel hypotheses about the
molecular genetic structure of working memory within the flexible RDoC
framework. The study PI is an investigator and data analyst for COGENT, which
was formed to conduct large-scale GWAS of cognitive traits implicated in
schizophrenia such as general cognitive ability, processing speed, working
memory, and episodic memory. At present, we have individual-level GWAS data on
24,000 participants who have 100,000 working memory data points collected for
analysis. To our knowledge, this would be the largest and most well powered GWAS
of working memory to date. The following interrelated specific aims are
proposed:

**Aim 1. Conduct a common factor GWAS of working memory**.

*Method*. Genomic structural equation modeling (Genomic SEM) will be used to
ascertain the joint genetic architecture of narrow working memory facets to
identify loci with effects on a genetically-derived general latent working
memory factor ("*Gwm*"). *Hypothesis*. With sample sizes in the tens of
thousands, the first replicable GWAS hits for working memory are expected to
emerge.

**Aim 2. Determine the extent of shared genetic architecture.**

*Method*. At genome-wide scale, linkage disequilibrium score regression (LDSC)
and genomic restricted maximum likelihood (GREML) estimation will be used to
determine the range of pleiotropy between working memory and correlated CNS
phenotypes. *Hypothesis*. Widespread coheritability is expected to emerge, and
*Gwm* will likely outperform narrower working memory domains and individual
working memory subtests.

**Aim 3: Isolate the strongest causal loci**.

*Method*. At the molecular level, Mendelian randomization (MR) will be used to
determine the direction of causality underlying pleiotropy between working
memory and correlated CNS phenotypes. *Hypothesis*. *Gwm* hits are expected to
exhibit strong causal effects from working memory to ADHD and schizophrenia, and
moderate causal effects to other CNS phenotypes.

**Aim 4. Map key causal biological mechanisms**.

*Method*. To prioritize *Gwm* loci for follow-up studies, functional mapping and
annotation of genetic associations (FUMA) will be used to characterize the
biological function of causal loci. *Hypothesis*. Elucidating the biology of
working memory is expected to be streamlined by narrowing the search space to
causal CNS mechanisms.

### B.1.a Have the major goals changed since the initial competing award or previous report?

No

## B.2 What was accomplished under these goals?

Year 1 was fraught with several challenges, some obvious such as the pandemic,
which limited access to our labs at USC for several months. and some less not
the least of which being the pandemic. During Year 1, we focused on designing
and finalizing the intervention that we will use in the RCT. The development of
the intervention involved many steps. We first looked to the Aim 1 survey
findings, which we completed in Year 1 (N = 527), to inform the intervention
content. For example, we included pregaming norms from the Aim 1 sample to
include in a section of the intervention on normative feedback. Next, we named
the intervention PACE (Pregaming Awareness in College Environments) to make it
accessible and recognizable to our targeted student population. We drafted a
storyboard of the full intervention (including the script for the narrator to
walk the participants through the intervention) and vetted all content with
college students, digital health experts, and our consultants familiar with
efficacious college student drinking interventions and app-based interventions.
We refined the storyboard based on feedback and began filming the narrator audio
and video. We added closed-captioning to the videos and programmed the
storyboard into a secure mobile-friendly website (<https://pregaming.net/>). Our
team, consultants, and students beta-tested the intervention on mobile phones
and provided feedback on usability. The web developers are in the final stages
of the website, which will be ready for a final test with participants in the
fall of 2020 (end of Year 2 and the start of Year 3).

### High-Performance Computing

Computational research often requires resources that exceed those of a personal
laptop or desktop computer. High-performance computing aggregates the resources
from individual computers (known as nodes) into a cluster that works together to
perform advanced, specialized computing jobs. The Center for Advanced Research
Computing launched its new high-performance computing cluster, Discovery, in
August 2020. The new cluster marks a significant upgrade to the CARC's
cyberinfrastructure, and the first step in a major, user-focused overhaul of the
program. The Discovery cluster consists of 2 shared login nodes and a total of
around 11,000 CPU cores in around 500 compute nodes. The typical compute node
has dual 8 to 16 core processors and resides on a 56 gigabit FDR InfiniBand
backbone. Nearly 100 GB of storage space is provided for each user to store
important code and configuration files in their home directories, and the
BeeGFS/ZFS parallel project file system has a capacity of 8.4 PB of usable
space, with a default storage quota of 10 TB per Principal Investigator across
their projects. A rapidly growing fleet of state-of-the art multicore compute
nodes is available for data-intensive research jobs, backed by two 800 TB,
high-throughput scratch file systems. A low-latency, high-bandwidth InfiniBand
network fabric facilitates intense research workloads on large data sets.
Discovery includes an array of scientific software packages, both licensed and
open source, for engineering, molecular simulation, and computational chemistry.
Researchers can also install software packages or develop their own code within
their project's allotted storage. Discovery is free to use for all USC faculty,
research staff, and graduate students (with the approval of their faculty
advisor). For information on how to set up your account and start using the
cluster, see the Accounts page. If you anticipate working with legally
restricted or sensitive data (e.g., HIPAA-, FERPA-, or CUI-regulated data), see
the Secure Computing page. Researchers also have the option of participating in
the Condo Cluster Program (CCP). The CCP allows researchers to reserve computing
resources for dedicated use by their projects, giving them the flexibility of
having their own cluster without the responsibility of purchasing hardware or
performing maintenance. For more information on the CCP, see the Condo Cluster
Program page.

### The Penn Letter N-Back Test

The Penn Letter N-Back Test measures working memory (WM), the ability to keep
and refresh goal-related information. Participants attend to a continual series
of letters that flash on the screen (one at a time) and press the spacebar
according to three different rules (called the 0-back, 1-back, and 2-back).
During the 0-back condition, the participant must simply respond to a currently
present target ("X"). During the 1-back condition, he/she must press the
spacebar when the letter on the screen is the same as the previous letter.
During the 2-back condition, he/she must press the spacebar when the letter on
the screen is the same as the letter before the previous letter (i.e. 2 letters
back). In all trials, the inter-stimulus interval (ISI) is 2.5 seconds, and the
stimuli (letters) themselves are presented for 0.5 seconds each. The participant
practices all three principles before testing.

## B.3 Competitive Revisions/Administrative Supplements

### For this reporting period, is there one or more Revision/Supplement associated with this award for which reporting is required?

No

## B.4 What opportunities for training and professional development has the project provided?

Many many. Reproducible research. R. Rmarkdown. GitHub. Git.

## B.5 How have the results been disseminated to communities of interest?

Neuropsychopharmacology

## B.6 What do you plan to do during the next reporting period to accomplish the goals?

```{=html}
<!-- 
\begin{wrapfigure}{R}{0.7\textwidth}

\hfill{}\includegraphics[width=.7\textwidth]{image.pdf} 

\caption{My Flowchart}\label{fig:unnamed-chunk-2}
\end{wrapfigure}
-->
```
```{r echo = F, warning = F, message = F, fig.width=7, fig.height = 6, out.width = ".3\\textwidth", fig.cap = "Important scatterplot", fig.align="right", wrapfigure = list("R", .3)}
plot(mpg ~ hp, data = mtcars)
```

### Timeline

There are lots of good examples of `R`-based Gantt charts to be found by clever
Googling. For displaying progress with sidebar annotations by aim, I
particularly like
[\underline{this}](https://datascienceplus.com/visualize-your-cvs-timeline-with-r-gantt-style/)
example from the [\underline{lares}](https://github.com/laresbernardo/lares)
package.
