---
title: "RPPR 2021"
subtitle: "GWAS of Working Memory"
author: "Joey Trampush PhD"
date: "`r Sys.Date()`"
output: tint::tintPdf
bibliography: bib/master.bib
link-citations: yes
institute: University of Southern California
---

```{r setup, include=FALSE}
library(tint)
# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tint'))
options(htmltools.dir.version = FALSE)
```

# Accomplishments

## B.1 What are the major goals of the project?

**SPECIFIC AIMS**

**Working Memory**, the most extensively studied of the executive functions, is
the cognitive system responsible for the transient holding and processing of
information in conscious awareness just long enough to use it for problem
solving. Limited working memory faculties are a fundamental aspect of the
cognitive impairments prevalent in ADHD, schizophrenia, and many other CNS
disorders. More than half of the variability in working memory capacity from
person to person can be sourced back to inherited genetic factors. Because it is
so highly heritable and diagnostically unbounded, the NIMH Research Domain
Criteria (RDoC) initiative has prioritized working memory as a core paradigmatic
construct to accelerate gene discovery research. Yet, to date, very few
genome-wide significant findings have been reported for working memory, and none
have been replicated.

As the discovery of loci reliably associated with working memory is key to
understanding neuropathophysiology, our goal is to fast-track the identification
and localization of molecular genetic signatures of working memory, and then to
uncover the underlying causal molecular and biological mechanisms linking
working memory deficits to genetically correlated disorders. GWAS (genome-wide
association study) is a powerful approach to identify molecular signatures of
complex traits if adequate sample sizes and proper statistical procedures are
employed. We will utilize existing genomic and phenotypic data from the
Cognitive Genomics Consortium (COGENT) to test novel hypotheses about the
molecular genetic structure of working memory within the flexible RDoC
framework. The study PI is an investigator and data analyst for COGENT, which
was formed to conduct large-scale GWAS of cognitive traits implicated in
schizophrenia such as general cognitive ability, processing speed, working
memory, and episodic memory. At present, we have individual-level GWAS data on
24,000 participants who have 100,000 working memory datapoints collected for
analysis. To our knowledge, this would be the largest and most well powered GWAS
of working memory to date. The following interrelated specific aims are
proposed:

**Aim 1. Conduct a common factor GWAS of working memory.**

Method. Genomic structural equation modeling (Genomic SEM) will be used to
ascertain the joint genetic architecture of narrow working memory facets to
identify loci with effects on a genetically-derived general latent working
memory factor ("Gwm"). Hypothesis. With sample sizes in the tens of thousands,
the first replicable GWAS hits for working memory are expected to emerge.

**Aim 2. Determine the extent of shared genetic architecture.**

Method. At genome-wide scale, linkage disequilibrium score regression (LDSC) and
genomic restricted maximum likelihood (GREML) estimation will be used to
determine the range of pleiotropy between working memory and correlated CNS
phenotypes. Hypothesis. Widespread coheritability is expected to emerge, and Gwm
will likely outperform narrower working memory domains and individual working
memory subtests.

**Aim 3: Isolate the strongest causal loci.**

Method. At the molecular level, Mendelian randomization (MR) will be used to
determine the direction of causality underlying pleiotropy between working
memory and correlated CNS phenotypes. Hypothesis. Gwm hits are expected to
exhibit strong causal effects from working memory to ADHD and schizophrenia, and
moderate causal effects to other CNS phenotypes.

**Aim 4. Map key causal biological mechanisms.**

Method. To prioritize Gwm loci for follow-up studies, functional mapping and
annotation of genetic associations (FUMA) will be used to characterize the
biological function of causal loci. Hypothesis. Elucidating the biology of
working memory is expected to be streamlined by narrowing the search space to
causal CNS mechanisms.

By deciphering the causal mechanisms through which allelic variation either
perturbs or protects the working memory system, which in turn can disrupt or
support the development and function of neural systems underlying working memory
across the lifespan, we can leverage the data to enhance theoretical and
neurobiological models of working memory more solidly grounded in molecular
genetics for a wide range of applications, which merits in-depth investigation.

## 1.a Have the major goals changed since the initial competing award or previous report?

[x] No

## B.2 What was accomplished under these goals?

For this reporting period describe: 1) major activities; 2) specific objectives;
3) significant results, including major findings, developments, or conclusions
(both positive and negative); and 4) key outcomes or other achievements. Include
a discussion of stated goals not met. As the project progresses, the emphasis in
reporting in this section should shift from reporting activities to reporting
accomplishments.

DHHS Specific Language Goals are equivalent to specific aims. In the response,
emphasize the significance of the findings to the scientific field. Include the
approaches taken to ensure robust and unbiased results. For most NIH awards the
response should not exceed 2 pages.

## Upload accomplishments [**Add Attachment**]

## B.3 Competitive Revisions/Administrative Supplements

For this reporting period, is there one or more Revision/Supplement associated
with this award for which reporting is required? Yes [x] No

If yes, identify the Revision(s)/Supplement(s) by grant number (e.g.,
3R01CA098765-01S1) or title and describe the specific aims and accomplishments
for each Revision/Supplement funded during this reporting period. Include any
supplements to promote diversity or re-entry, or other similar supplements to
support addition of an individual or a discrete project.

Revision/Supplement

or Revision/Supplement Title Total remaining allowed limit is 255 characters.
Describe the specific aims for this Revision/Supplement below (Limit is 700
characters or approximately 1/4 of a page.) Total remaining allowed limit is 700
characters. Describe the accomplishments for this Revision/Supplement below
(Limit is 700 characters or approximately 1/4 of a page.) Total remaining
allowed limit is 700 characters.

No items found. Revision/Supplement \# Revision/Supplement Title Specific Aims
Accomplishments Action Nothing found to display. ---

## B.4 What opportunities for training and professional development has the project provided?

If the research is not intended to provide training and professional development
opportunities or there is nothing significant to report during this reporting
period, state "Nothing to Report."

Describe opportunities for training and professional development provided to
anyone who worked on the project or anyone who was involved in the activities
supported by the project. "Training" activities are those in which individuals
with advanced professional skills and experience assist others in attaining
greater proficiency. Training activities may include, for example, courses or
one-on-one work with a mentor. "Professional development" activities result in
increased knowledge or skill in one's area of expertise and may include
workshops, conferences, seminars, study groups, and individual study. Include
participation in conferences, workshops, and seminars not listed under major
activities.

DHHS Specific Language For all projects reporting graduate student and/or
postdoctoral participants in Section D. Participant, grantees are encouraged to
describe the use of Individual Development Plans (IDPs) for those participants.A
Do not include the actual IDP; instead include information to document that IDPs
are used to help manage the training for those individuals.

DHHS Specific Language For T, F, K, R25, R13, D43 and other awards or award
components designed to provide training and professional development
opportunities, a response is required. Do not reiterate what is reported under
Accomplishments. Limit the response to this reporting period.

Nothing to Report or [x] **Upload Description and Diversity Report, as
applicable**

## High-Performance Computing

Computational research often requires resources that exceed those of a personal
laptop or desktop computer. High-performance computing aggregates the resources
from individual computers (known as nodes) into a cluster that works together to
perform advanced, specialized computing jobs. The Center for Advanced Research
Computing launched its new high-performance computing cluster, Discovery, in
August 2020. The new cluster marks a significant upgrade to the CARC's
cyberinfrastructure, and the first step in a major, user-focused overhaul of the
program. The Discovery cluster consists of 2 shared login nodes and a total of
around 11,000 CPU cores in around 500 compute nodes. The typical compute node
has dual 8 to 16 core processors and resides on a 56 gigabit FDR InfiniBand
backbone. Nearly 100 GB of storage space is provided for each user to store
important code and configuration files in their home directories, and the
BeeGFS/ZFS parallel project file system has a capacity of 8.4 PB of usable
space, with a default storage quota of 10 TB per Principal Investigator across
their projects. A rapidly growing fleet of state-of-the art multicore compute
nodes is available for data-intensive research jobs, backed by two 800 TB,
high-throughput scratch file systems. A low-latency, high-bandwidth InfiniBand
network fabric facilitates intense research workloads on large data sets.
Discovery includes an array of scientific software packages, both licensed and
open source, for engineering, molecular simulation, and computational chemistry.
Researchers can also install software packages or develop their own code within
their project's allotted storage. Discovery is free to use for all USC faculty,
research staff, and graduate students (with the approval of their faculty
advisor). For information on how to set up your account and start using the
cluster, see the Accounts page. If you anticipate working with legally
restricted or sensitive data (e.g., HIPAA-, FERPA-, or CUI-regulated data), see
the Secure Computing page. Researchers also have the option of participating in
the Condo Cluster Program (CCP). The CCP allows researchers to reserve computing
resources for dedicated use by their projects, giving them the flexibility of
having their own cluster without the responsibility of purchasing hardware or
performing maintenance. For more information on the CCP, see the Condo Cluster
Program page.

## The Penn Letter N-Back Test

The Penn Letter N-Back Test measures working memory (WM), the ability to keep
and refresh goal-related information. Participants attend to a continual series
of letters that flash on the screen (one at a time) and press the spacebar
according to three different rules (called the 0-back, 1-back, and 2-back).
During the 0-back condition, the participant must simply respond to a currently
present target ("X"). During the 1-back condition, he/she must press the
spacebar when the letter on the screen is the same as the previous letter.
During the 2-back condition, he/she must press the spacebar when the letter on
the screen is the same as the letter before the previous letter (i.e. 2 letters
back). In all trials, the inter-stimulus interval (ISI) is 2.5 seconds, and the
stimuli (letters) themselves are presented for 0.5 seconds each. The participant
practices all three principles before testing.

## B.5 How have the results been disseminated to communities of interest?

Describe how the results have been disseminated to communities of interest.
Include any outreach activities that have been undertaken to reach members of
communities who are not usually aware of these research activities, for the
purpose of enhancing public understanding and increasing interest in learning
and careers in science, technology, and the humanities.

DHHS Specific Language Reporting the routine dissemination of information (e.g.,
websites, press releases) is not required. For awards not designed to
disseminate information to the public or conduct similar outreach activities, a
response is not required and the grantee should select "Nothing to Report". A
detailed response is only required for awards or award components that are
designed to disseminate information to the public or conduct similar outreach
activities. Note that scientific publications and the sharing of research
sources will be reported under Products.

Nothing to Report

**or enter response below (NIH recommended length is up to 1 page.** Limit is
8000 characters or approximately 3 pages.)

Total remaining allowed limit is 8000 characters.

## B.6 What do you plan to do during the next reporting period to accomplish the goals?

Describe briefly what you plan to do during the next reporting period to
accomplish the goals and objectives.

DHHS Specific Language Discuss efforts to ensure that the approach is
scientifically rigorous and results are robust and unbiased. Remember that
significant changes in objectives and scope require prior approval of the agency
(e.g., NIH Grants Policy Statement, 8.1.2.).

DHHS Specific Language Include any important modifications to the original
plans. Provide a scientific justification for any changes involving research
with human subjects or vertebrate animals. A detailed description of such
changes must be provided under Changes.

**Enter response below (NIH recommended length is up to 1 page. Limit is 8000
characters or approximately 3 pages.)** Total remaining allowed limit is 8000
characters.

> insert

> end

# Products

## publications, conference papers, and presentations

## website(s) or other Internet site(s)

## technologies or techniques

## inventions, patent applications, and/or licenses

## other products, such as data or databases, physical collections, audio or video products, software, models, educational aids or curricula, instruments or equipment, research material, interventions (e.g., clinical or educational), or new business creation.

# Participants and Other Collaborating Organizations

Amanda Law Jack Bowden

# Impact

Zero

# Changes/Problems (not required for Final or Interim RPPR)

## Changes in approach and reasons for change

## Actual or anticipated problems or delays and actions or plans to resolve them

## Changes that have a significant impact on expenditures

## Significant changes in use or care of human subjects, vertebrate animals, biohazards, and/or select agents

# Budgetary Information (not required for Final or Interim RPPR)

N/A

## Project Outcomes (only required on Final and Interim RPPR)

## Concise summary of the outcomes or findings of the award, written for the general public in clear and comprehensible language, without including any proprietary, confidential information or trade secrets

Note: project outcome information will be made public in NIH RePORTER Link to
External Site

# Before We Get Started...

```{r fig-nocap-margin-first, fig.margin=TRUE, fig.width=4, fig.height=4, cache=TRUE, echo=FALSE}
library(ggplot2)
ggplot(mtcars, aes(wt, mpg)) + geom_point(size=3, aes(colour=factor(cyl))) + theme(legend.position="none")
```

[tint](http://github.com/eddelbuettel/tint) is straightforward mix of the (html
and pdf parts of the) excellent [tufte](https://github.com/rstudio/tufte)
package by JJ and Yihui, mixed with the [Roboto
Condensed](https://fonts.google.com/specimen/Roboto+Condensed) font use and
color scheme proposed by [envisioned
css](https://github.com/nogginfuel/envisioned-css) plus minor style changes such
as removal of *italics*---but otherwise true to the wonderful
[tufte](https://github.com/rstudio/tufte) package for R---all baked together
into a small package providing another template.

NB: This `tintBook` variant is for book-length document in pdf; there are is
also short `tintPdf` variant suitable for handouts or articles along with a
similar `tintHtml` variant for shorter web-based documents. There is currently
no book-length html variant as we find all current html / css / js offerings for
this unsatisfactory. Also note that we use the exact same input as for the other
two formats and have not yet customized this for a book.

We support *italic* aka *em* and **strong** annotations for text, as well as
`code` snippets.

The package name is a standard package naming recursion: *tint is not tufte*.

The remainder of the [tufte](https://github.com/rstudio/tufte) skeleton document
follows as is, with only marginal changes to refer to this package for code.

# Introduction

The Tufte handout style is a style that Edward Tufte uses in his books and
handouts. Tufte's style is known for its extensive use of sidenotes, tight
integration of graphics with text, and well-set typography. This style has been
implemented in LaTeX and HTML/CSS[^1], respectively. We have ported both
implementations into the [**tufte** package](https://github.com/rstudio/tufte).
If you want LaTeX/PDF output, you may use the `tufte_handout` format for
handouts, and `tufte_book` for books. For HTML output, use `tufte_html`. These
formats can be either specified in the YAML metadata at the beginning of an R
Markdown document (see an example below), or passed to the `rmarkdown::render()`
function. See @R-rmarkdown for more information about **rmarkdown**.

[^1]: See Github repositories
    [tufte-latex](https://github.com/tufte-latex/tufte-latex) and
    [tufte-css](https://github.com/edwardtufte/tufte-css)

``` {.yaml}
---
title: "An Example Using the Tufte Style"
author: "John Smith"
output:
  tufte::tufte_handout: default
  tufte::tufte_html: default
---
```

There are two goals of this package:

1.  To produce both PDF and HTML output with similar styles from the same R
    Markdown document;
2.  To provide simple syntax to write elements of the Tufte style such as side
    notes and margin figures, e.g. when you want a margin figure, all you need
    to do is the chunk option `fig.margin = TRUE`, and we will take care of the
    deails for you, so you never need to think about
    `\begin{marginfigure} \end{marginfigure}` or
    `<span class="marginfigure"> </span>`; the LaTeX and HTML code under the
    hood may be complicated, but you never need to learn or write such code.

If you have any feature requests or find bugs in **tufte**, please do not
hesitate to file them to <https://github.com/rstudio/tufte/issues>. For general
questions, you may ask them on StackOverflow:
<http://stackoverflow.com/tags/rmarkdown>.

# Headings

This style provides first and second-level headings (that is, `#` and `##`),
demonstrated in the next section. You may get unexpected output if you try to
use `###` and smaller headings.

`r newthought('In his later books')`[^2], Tufte starts each section with a bit
of vertical space, a non-indented paragraph, and sets the first few words of the
sentence in small caps. To accomplish this using this style, call the
`newthought()` function in **tufte** in an *inline R expression* `` `r ` `` as
demonstrated at the beginning of this paragraph.[^3]

[^2]: [Beautiful Evidence](https://www.edwardtufte.com/tufte/books_be)

[^3]: Note you should not assume **tufte** has been attached to your R session.
    You should either `library(tufte)` in your R Markdown document before you
    call `newthought()`, or use `tint::newthought()`.

# Figures

## Margin Figures

Images and graphics play an integral role in Tufte's work. To place figures in
the margin you can use the **knitr** chunk option `fig.margin = TRUE`. For
example:

```{r fig-margin, fig.margin = TRUE, fig.cap = "MPG vs horsepower, colored by transmission.", fig.width=3.5, fig.height=3.5, cache=TRUE}
library(ggplot2)
mtcars2 <- mtcars
mtcars2$am <- factor(
  mtcars$am, labels = c('automatic', 'manual')
)
ggplot(mtcars2, aes(hp, mpg, color = am)) +
  geom_point() + geom_smooth() +
  theme(legend.position = 'bottom')
```

Note the use of the `fig.cap` chunk option to provide a figure caption. You can
adjust the proportions of figures using the `fig.width` and `fig.height` chunk
options. These are specified in inches, and will be automatically scaled down to
fit within the handout margin.

## Arbitrary Margin Content

In fact, you can include anything in the margin using the **knitr** engine named
`marginfigure`. Unlike R code chunks ```` ```{r} ````, you write a chunk
starting with ```` ```{marginfigure} ```` instead, then put the content in the
chunk. See an example on the right about the first fundamental theorem of
calculus.

```{marginfigure}
We know from _the first fundamental theorem of calculus_ that for $x$ in $[a, b]$:
$$\frac{d}{dx}\left( \int_{a}^{x} f(u)\,du\right)=f(x).$$
```

For the sake of portability between LaTeX and HTML, you should keep the margin
content as simple as possible (syntax-wise) in the `marginefigure` blocks. You
may use simple Markdown syntax like `**bold**` and `_italic_` text, but please
refrain from using footnotes, citations, or block-level elements (e.g.
blockquotes and lists) there.

## Full Width Figures

You can arrange for figures to span across the entire page by using the chunk
option `fig.fullwidth = TRUE`.

```{r fig-fullwidth, fig.width = 10, fig.height = 2, fig.fullwidth = TRUE, fig.cap = "A full width figure.", warning=FALSE, cache=TRUE}
ggplot(diamonds, aes(carat, price)) + geom_smooth() +
  facet_grid(~ cut)
```

Other chunk options related to figures can still be used, such as `fig.width`,
`fig.cap`, `out.width`, and so on. For full width figures, usually `fig.width`
is large and `fig.height` is small. In the above example, the plot size is
$10 \times 2$.

## Main Column Figures

Besides margin and full width figures, you can of course also include figures
constrained to the main column. This is the default type of figures in the
LaTeX/HTML output.

```{r fig-main, fig.cap = "A figure in the main column.", cache=TRUE}
ggplot(diamonds, aes(cut, price)) + geom_boxplot()
```

# Sidenotes

One of the most prominent and distinctive features of this style is the
extensive use of sidenotes. There is a wide margin to provide ample room for
sidenotes and small figures. Any use of a footnote will automatically be
converted to a sidenote. [^4]

[^4]: This is a sidenote that was entered using a footnote.

If you'd like to place ancillary information in the margin without the sidenote
mark (the superscript number), you can use the `margin_note()` function from
**tufte** in an inline R expression.
`r margin_note("This is a margin note.  Notice that there is no number preceding the note.")`
This function does not process the text with Pandoc, so Markdown syntax will not
work here. If you need to write anything in Markdown syntax, please use the
`marginfigure` block described previously.

# References

References can be displayed as margin notes for HTML output. For example, we can
cite R here [@R-base]. To enable this feature, you must set
`link-citations: yes` in the YAML metadata, and the version of `pandoc-citeproc`
should be at least 0.7.2. You can always install your own version of Pandoc from
<http://pandoc.org/installing.html> if the version is not sufficient. To check
the version of `pandoc-citeproc` in your system, you may run this in R:

```{r eval=FALSE}
system2('pandoc-citeproc', '--version')
```

If your version of `pandoc-citeproc` is too low, or you did not set
`link-citations: yes` in YAML, references in the HTML output will be placed at
the end of the output document.

# Tables

You can use the `kable()` function from the **knitr** package to format tables
that integrate well with the rest of the Tufte handout style. The table captions
are placed in the margin like figures in the HTML output.

```{r}
knitr::kable(
  mtcars[1:6, 1:6], caption = 'A subset of mtcars.'
)
```

# Block Quotes

We know from the Markdown syntax that paragraphs that start with `>` are
converted to block quotes. If you want to add a right-aligned footer for the
quote, you may use the function `quote_footer()` from **tufte** in an inline R
expression. Here is an example:

> "If it weren't for my lawyer, I'd still be in prison. It went a lot faster
> with two people digging."
>
> `r tint::quote_footer('--- Joe Martin')`

Without using `quote_footer()`, it looks like this (the second line is just a
normal paragraph):

> "Great people talk about ideas, average people talk about things, and small
> people talk about wine."
>
> --- Fran Lebowitz

# Responsiveness

The HTML page is responsive in the sense that when the page width is smaller
than 760px, sidenotes and margin notes will be hidden by default. For sidenotes,
you can click their numbers (the superscripts) to toggle their visibility. For
margin notes, you may click the circled plus signs to toggle visibility.

# More Examples

The rest of this document consists of a few test cases to make sure everything
still works well in slightly more complicated scenarios. First we generate two
plots in one figure environment with the chunk option `fig.show = 'hold'`:

```{r fig-two-together, fig.cap="Two plots in one figure environment.", fig.show='hold', cache=TRUE}
p <- ggplot(mtcars2, aes(hp, mpg, color = am)) +
  geom_point()
p
p + geom_smooth()
```

Then two plots in separate figure environments (the code is identical to the
previous code chunk, but the chunk option is the default `fig.show = 'asis'`
now):

```{r fig-two-separate, ref.label='fig-two-together', fig.cap=sprintf("Two plots in separate figure environments (the %s plot).", c("first", "second")), cache=TRUE}
```

You may have noticed that the two figures have different captions, and that is
because we used a character vector of length 2 for the chunk option `fig.cap`
(something like `fig.cap = c('first plot', 'second plot')`).

Next we show multiple plots in margin figures. Similarly, two plots in the same
figure environment in the margin:

```{r fig-margin-together, fig.margin=TRUE, fig.show='hold', fig.cap="Two plots in one figure environment in the margin.", fig.width=3.5, fig.height=2.5, cache=TRUE}
p
p + geom_smooth(method = 'lm')
```

Then two plots from the same code chunk placed in different figure environments:

```{r fig-margin-separate, fig.margin=TRUE, fig.cap=sprintf("Two plots in separate figure environments in the margin (the %s plot).", c("first", "second")), fig.width=3.5, fig.height=2.5, cache=TRUE}
knitr::kable(head(iris, 15))
p
knitr::kable(head(iris, 12))
p + geom_smooth(method = 'lm')
knitr::kable(head(iris, 5))
```

We blended some tables in the above code chunk only as *placeholders* to make
sure there is enough vertical space among the margin figures, otherwise they
will be stacked tightly together. For a practical document, you should not
insert too many margin figures consecutively and make the margin crowded.

You do not have to assign captions to figures. We show three figures with no
captions below in the margin, in the main column, and in full width,
respectively.

```{r fig-nocap-margin, fig.margin=TRUE, fig.width=3.5, fig.height=2, cache=TRUE}
# a boxplot of weight vs transmission; this figure
# will be placed in the margin
ggplot(mtcars2, aes(am, wt)) + geom_boxplot() +
  coord_flip()
```

```{r fig-nocap-main, cache=TRUE}
# a figure in the main column
p <- ggplot(mtcars, aes(wt, hp)) + geom_point()
p
```

```{r fig-nocap-fullwidth, fig.fullwidth=TRUE, fig.width=10, fig.height=3, cache=TRUE}
# a fullwidth figure
p + geom_smooth(method = 'lm') + facet_grid(~ gear)
```

# Some Notes on Tufte CSS

There are a few other things in Tufte CSS that we have not mentioned so far. If
you prefer `r sans_serif('sans-serif fonts')`, use the function `sans_serif()`
in **tufte**. For epigraphs, you may use a pair of underscores to make the
paragraph italic in a block quote, e.g.

> *I can win an argument on any topic, against any opponent. People know this,
> and steer clear of me at parties. Often, as a sign of their great respect,
> they don't even invite me.*
>
> `r quote_footer('--- Dave Barry')`

We hope you will enjoy the simplicity of R Markdown and this R package, and we
sincerely thank the authors of the Tufte-CSS and Tufte-LaTeX projects for
developing the beautiful CSS and LaTeX classes. Our **tufte** package would not
have been possible without their heavy lifting.

To see the R Markdown source of this example document, you may follow [this link
to
Github](https://github.com/eddelbuettel/tint/blob/master/inst/rmarkdown/templates/pdf/skeleton/skeleton.Rmd),
use the wizard in RStudio IDE
(`File -> New File -> R Markdown -> From Template`), or open the Rmd file in the
package:

```{r eval=FALSE}
file.edit(
  tint:::template_resources(
    'tint', '..', 'skeleton', 'skeleton.Rmd'
  )
)
```

```{r bib, include=FALSE}
# create a bib file for the R packages used in this document
knitr::write_bib(c('base', 'rmarkdown'), file = 'skeleton.bib')
```
