---
title: "Using hrbrthemes in PDFs"
output: hrbrthemes::ipsum_pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  dev = "cairo_pdf")
```

Captain John Sheridan: You know, I just had a thought. You've been back and
forth to your world so many times since you got here. How do I know you're the
same Vorlon? Inside that encounter suit you could be anyone. Kosh Naranek: I
have *always* been here. Captain John Sheridan: Oh, yeah? You said that about me
too. Kosh Naranek: Yes. [starts to walk away] Captain John Sheridan: I really
*hate* it when you do that. Kosh Naranek: [turns around] Good! Delenn: I am
Grey. I stand between the candle and the star. We are Grey. We stand between the
darkness and the light.

Lt. Corwin: Do we trust no-one then? Cmdr. Susan Ivanova: No, trust Ivanova,
trust yourself, anybody else, shoot'em. Ta'Lon: Congratulations citizen G'Kar.
You are now a religious icon. Susan Ivanova: So the next time we find out where
the Shadows plan to strike, we can mine the area, and as soon as they come out
of hyperspace... Citizen G'Kar: Then, as you so concisely say, Boom!

## Section Title

G'Kar: It is said that the future is always born in pain. The history of war is
the history of pain. If we are wise, what is born of that pain matures into the
promise of a better world, because we learn that we can no longer afford the
mistakes of the past. Dr. Stephen Franklin: Doesn't anyone listen to one damn
word I say? Captain John Sheridan: No surrender, no retreat. [Opening narration,
season 4] Lennier: It was the year of fire, Zack Allan: The year of destruction,
Citizen G'Kar: The year we took back what was ours. Lyta Alexander: It was the
year of rebirth, Ambassador Vir Cotto: The year of great sadness, Marcus Cole:
The year of pain, Delenn: And a year of joy. Ambassador Londo Mollari: It was a
new age. Dr. Stephen Franklin: It was the end of history. Susan Ivanova: It was
the year everything changed. Michael Garibaldi: The year is 2261. Captain John
Sheridan: The place, Babylon 5. Sinclair: They say God works in mysterious ways.
Michael Garibaldi: Maybe so, but He's a con-man compared to the Vorlon.

# Purpose

The purpose of this template is to show what you need to do in order to use
Arial Narrow or Roboto Condensed in PDF output documents.

The main consideration is the use of `Cairo` and --- specifically -- the
`cairo_pdf` output device. It is best to set that in the `knit` chunk defaults
(as seen in the `setup` block) but you can also do that in the individual chunks
as has also been demonstrated in the "Roboto Condensed" chunk.

Make sure to include the library calls as they are here.

```{r message = FALSE}
library(hrbrthemes)
library(ggplot2)
library(Cairo)
library(extrafont)

extrafont::loadfonts()
```

### Arial Narrow

```{r}
ggplot(mtcars, aes(wt, mpg)) +
  geom_point() +
  labs(title="Arial Narrow", subtitle="This is a subtitle") +
  theme_ipsum()
```

### Roboto Condensed

```{r dev = "cairo_pdf"}
ggplot(mtcars, aes(wt, mpg)) +
  geom_point() +
  labs(title="Roboto Condensed", subtitle="This is a subtitle") +
  theme_ipsum_rc()
```
