---
title: "GenomicSEM Common Factor GWAS Analysis"
output: github_document
---

```{bash}
# transfer data using cli
rsync -avh /Users/joey/gwm.flow/reference.1000G.maf.0.005.txt.gz trampush@hpc-transfer2.usc.edu:/project/trampush_325/gwm
rsync -avh /Users/joey/gwm.flow/script-gsem-commonFactor.R trampush@hpc-transfer2.usc.edu:/project/trampush_325/gwm
scp ~/gwm.flow/scz_withRS.txt trampush@discovery.usc.edu:/project/trampush_325
## symlinks
ln -s /project/trampush_325/gwm/genotypes/gwas_sumstats/*.txt /project/trampush_325/gwm
```

```{bash}
### Starting R
module spider r
module spider r/4.1.0

### Testing your job
salloc --ntasks=4 --time=1:00:00 --account=<account_id>

### To run R interactively on a compute node, first use Slurm's salloccommand to reserve resources on a node:
## for GenomicSEM might have to ask for same number of cores as traits -- seems to be a bug
salloc --ntasks=4 --time=2:00:00 --cpus-per-task=8 --mem=16GB --account=trampush_325

# trying with this setup; might need a lot of memory!!
salloc --ntasks=1 --time=2:00:00 --cpus-per-task=4 --mem=32GB --account=trampush_325

### to run 4.1 
module load gcc/8.3.0 
module load openblas/0.3.8
module load r/4.1.0

# export library
# may or may not always need this
export R_LIBS_USER=/project/trampush_325/R/pkgs/4.1.0

### start R
R
```

```{r}
# Parallel prgramming wiht R
# Note: On CARC systems, 1 thread = 1 core = 1 logical CPU

```

```{r}
### Parallel programming with R
### submitting slurm job
### Note: On CARC systems, 1 thread = 1 core = 1 logical CPU

#!/bin/bash

#SBATCH --nodes=1
#SBATCH --ntasks=1
#SBATCH --cpus-per-task=8
#SBATCH --mem=16GB
#SBATCH --time=1:00:00
#SBATCH --account=trampush_325

module purge
module load gcc/8.3.0
module load openblas/0.3.8
module load r/4.1.0

Rscript --vanilla script.R
```

```{r}
# install packages from login node only
install.packages("devtools")
install.packages("githubinstall")
install.packages("tidyverse")
install.packages("data.table")
install.packages("here")
install.packages("rmarkdown")
install.packages("bookdown")
install.packages("languageserver")
install.packages("shiny")
install.packages("vroom")
devtools::install_github("GenomicSEM/GenomicSEM")
install.packages("renv")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, eval=FALSE}
# symlink GWAS summary stats
ln -shfi ~/gwm.flow/data/genotypes/w_hm3.noMHC.snplist ~/gwm.flow/w_hm3.noMHC.snplist 
ln -shfi ~/gwm.flow/data/genotypes/anx.sumstats.gz ~/gwm.flow/anx.sumstats.gz 
## to zcat and see inside zipped files
zcat < anx.sumstats.gz | wc -l
zcat < anx.sumstats.gz | more 
```

# Common Factor GWAS

```{r load}
# Start here
##read in the schizophrenia summary statistics using fread from the data.table R package##
library(tidyverse)
library(GenomicSEM)

library(data.table)
library(vroom)
require(here)
library(foreach)
library(future)
```

GenomicSEM for Common Factor GWAS A powerful extension of Genomic SEM is to run
a multivariate GWAS for which a common factor defined by genetic indicators is
regressed on a SNP. This allows for estimation of a set of summary statistics
for the common factor that represent the SNP effects on the common factor.
Estimation of a common factor GWAS proceeds in 5 steps. Note that estimation of
the GWAS models are all entirely independent of one another. This means that the
summary statistics used as input can be arbitrarily split up and submitted to
different jobs on a computing cluster to help with memory and run-time issues.
Please also note that the sumstats function requires that the variables be
listed in the same order as they are listed for the ldsc function.

## Step 1: Munge the summary statistics

As in the non-GWAS examples, the summary statistics need to be munged in order
to produce the S and V matrix that do not include the SNPs. All traits can be
munged at once as in the example below. In order to use our package to prepare
the summary statistics the munge function can be used to convert the summary
statistics to the format expected by LD-score Regression. The summary statistics
files input into the munge function at a minimum need to contain five pieces of
information:

1.  The rsid of the SNP.
2.  An A1 allele column, with A1 indicating the effect allele.
3.  An A2 allele column, with A2 indicating the non-effect allele.
4.  Either a logistic or continuous regression effect.
5.  The p-value associated with this effect.

The package will automatically rename the column based on commonly observed
names, but may return an error if the file contains trait-specific column
headers (e.g., RSID_SCHIZOPHRENIA). All traits can be munged at once as in the
example below. Here we use traits for Bipolar (2011), Major Depression (2018),
PTSD (2017), and Anxiety (2016) available for download from:
<http://www.med.unc.edu/pgc/results-and-downloads>. The Schizophrenia (2018)
summary statistics are available for download from:
<http://walters.psycm.cf.ac.uk/>.

The munge function takes 6 arguments:

1.  files: The name of the summary statistics files

```{r}
files = c(
   "adhd_pgc_2019.txt",
   "cogent_g_n38.txt",
   "GWAS_CP_all.txt")
```

1.  hm3: The name of the reference file. Here we use Hapmap 3 SNPs. This file
    can be obtained from
    <https://utexas.box.com/s/vkd36n197m8klbaio3yzoxsee6sxo11v>. Note that we
    previously used a reference file that removed the MHC (i.e., HLA) region,
    but upon further inspection found that the MHC region was also removed from
    the LD-score files used for estimating LD-Score regression in the following
    step, such that MHC is automatically excluded from analyses. You do NOT need
    to rerun analyses if you have been using the Hapmap3 SNPs with MHC removed
    as you should obtain equivalent results.

```{r}
hm3 = "w_hm3.snplist"
```

1.  trait.names: The trait names that will be used to name the saved files

```{r}
trait.names = c(
   "adhd",
   "cog",
   "cp") 
```

1.  N: The total sample sizes associated with the traits. Note that these should
    not be converted to effective sample sizes for case/control designs as the
    LDSC function already corrects for this by using the population and sample
    prevalence provided by the user. If the input contains SNP-specific sample
    sizes for each row, then the munge column will automatically use this N. If
    the input contains both SNP-specific and separate sample size columns for
    cases and controls, then the munge function will automatically combine these
    column to create a SNP-specific total sample size.

```{r}
N = c(
   55274,
   28445,
   257828)
```

1.  info.filter: The INFO filter. The package default is to filter to SNPs with
    INFO \> 0.9. If the summary statistics do not contain an INFO column the
    function can still be run, but results should be interpreted keeping in mind
    that this cleaning step was missed.

```{r}
info.filter = 0.9
```

1.  maf.filter: The MAF filter. The package default is to filter to SNPs with
    MAF \> 0.01.

```{r}
maf.filter = 0.01
```

We note that our munge function will produce slightly different results,
generally with less SNPs, then the munge function from the original LDSC
package. This is because, in addition to checking whether or not the A1 allele
matches A1/A2 in the ref file, our munge function also checks whether A2 matches
A1/A2 in the ref file. In addition, our munge function removes SNP with missing
MAFs in the summary statistics file.

```{r}
# awk these and changed file names
#daner_adhd_meta_filtered_NA_iPSYCH23_PGC11_sigPCs_woSEX_2ell6sd_EUR_Neff_70.meta.gz
#daner_cogent_38.filtered.results
MUNGEoutput <- munge(
   files = files,
   hm3 = hm3,
   trait.names = trait.names,
   N = N,
   info.filter = info.filter,
   maf.filter = maf.filter)

# save as file to look at it
readr::write_tsv(MUNGEoutput, "LDSCoutput.tsv")
```

## Step 2: Run multivariable LDSC

The second step is to run multivariable LD-Score regression to obtain the
genetic covariance (S) matrix and corresponding sampling covariance matrix (V).
This is achieved by running the ldsc function. We note that it would not be
appropriate in this case to take output from runs of the original LD score
regression to construct by hand the S and V matrix. This is because the sampling
covariances that occupy the off-diagonal elements of the V matrix could not be
filled in. The ldsc function takes 5 necessary arguments and some additional
optional arguments:

1.  traits: a vector of file names/paths to files which point to the munged
    sumstats.

```{r}
traits <- c(
   "cp.sumstats.gz",
   "cog.sumstats.gz",
   "adhd.sumstats.gz",
   "anx.sumstats.gz",
   "mdd.sumstats.gz",
   "scz.sumstats.gz"
   )
```

1.  sample.prev: A vector of sample prevalences of length equal to the number of
    traits. Sample prevalence is calculated as the number of cases over the
    total number of participants (i.e., cases + controls). Possible range = 0-1.
    If the trait is continuous, the values should equal NA.

```{r}
sample.prev <- c(
   NA,
   NA,
   0.37,
   0.32,
   0.35,
   0.39)
```

1.  population.prev: A vector of population prevalences. These estimates can be
    obtained from a number of sources, such as large scale epidemiological
    studies. Possible range = 0-1. Again, if the trait is continuous the values
    should equal NA.

```{r}
population.prev <- c(
   NA,
   NA,
   0.10,
   0.29,
   0.16,
   0.01)
```

1.  ld and wld: A folder of LD scores used as the independent variable in LDSC
    (ld) and LDSC weights (wld). These are typically the same folder, and in the
    original LD score package is called "eur_w\_ld_chr". We use the same LD
    scores and weights for our application, though the user can supply their own
    if desired. Weights for the european population used here can be obtained
    from by downloading the eur_w\_ld_chr folder (note that these are the same
    weights provided by the original developers of LDSC):
    <https://utexas.box.com/s/vkd36n197m8klbaio3yzoxsee6sxo11v>. If you receive
    an error when running ldsc of not being able to change your working
    directory please be sure that you have specified the correct file path to
    the LD scores. Please note that the LD scores must match the ancestral
    background of the summary statistics provided (i.e., European summary
    statistics with European LD scores). It is not currently possible to run
    LD-score regression with ad-mixed ancestral backgrounds. The following
    example takes summary statistics for the p-factor and runs multivariable LD
    score regression.

```{r}
ld <- "eur_w_ld_chr/"
```

1.  ld and wld: A folder of LD scores used as the independent variable in LDSC
    (ld) and LDSC weights (wld). These are typically the same folder, and in the
    original LD score package is called "eur_w\_ld_chr". We use the same LD
    scores and weights for our application, though the user can supply their own
    if desired. Weights for the european population used here can be obtained
    from by downloading the eur_w\_ld_chr folder (note that these are the same
    weights provided by the original developers of LDSC):
    <https://utexas.box.com/s/vkd36n197m8klbaio3yzoxsee6sxo11v>. If you receive
    an error when running ldsc of not being able to change your working
    directory please be sure that you have specified the correct file path to
    the LD scores. Please note that the LD scores must match the ancestral
    background of the summary statistics provided (i.e., European summary
    statistics with European LD scores). It is not currently possible to run
    LD-score regression with ad-mixed ancestral backgrounds. The following
    example takes summary statistics for the p-factor and runs multivariable LD
    score regression.

```{r}
wld <- "eur_w_ld_chr/"
```

1.  trait.names: An optional argument specifying the trait names. This allows
    for model specification using the trait names in later steps, which can be
    useful for keeping track of results when the number of traits becomes large.
    If this argument is not specified, the function will automatically name the
    traits in the general form V1-VX.

```{r}
trait.names <- c(
   "CP",
   "COG",
   "ADHD",
   "ANX",
   "MDD",
   "SCZ")
```

1.  ldsc.log: An optional argument specifying how you want to name the ldsc.log
    file. The default is to name the log file using the file names of the munged
    summary statistics.

```{r}
##ldsc.log <- "gsem.test.run.3.traits.gcog1.gcog2.adhd.log"
```

1.  stand: An optional argument specifying whether you want the ldsc to also
    output the genetic correlation matrix and its sampling covariance matrix.
    Default = FALSE.

```{r}
stand=TRUE
```

### Run `ldsc` analysis

```{r}
LDSCoutput <- ldsc(
   traits = traits, 
   sample.prev = sample.prev, 
   population.prev = population.prev, 
   ld = ld, 
   wld = wld, 
   trait.names = trait.names,
   stand = stand
   )

# save as file to look at it
readr::write_file(LDSCoutput, "LDSCoutput.file", append = FALSE)
```

## Step 3: Prepare Summary Statistics for GWAS

The sumstats function takes care of a few things prior to running multivariate
GWAS. In order to run a multivariate GWAS, the SNPs, and corresponding SNP
effects, need to be coded across phenotypes so that the same allele is the
reference allele in all cases. In all cases, coefficients and their SEs are then
further transformed such that they are scaled relative to unit-variance scaled
phenotypes. In order to appropriately perform this standardization for different
types of betas (e.g., logistic betas, odds-ratios), the sumstats function takes
a number of arguments outlined below. The sumstats function then combines these
newly scaled betas and SEs into a single, listwise deleted file. Note that the
OLS and LPM transformation require only a total sample size, p-value, and signed
effect column (e.g., beta or Z-statistic), and do not require a SE column.
Transformations for a case/control GWAS require both a SE column and either an
odds ratio or logistic beta. Confused about how to specify the arguments for
this function? See point 3 on the important resources and key information page.
Please also note that the sumstats function requires that the variables be
listed in the same order as they are listed for the ldsc function above. The
sumstats function takes 4 necessary arguments, along with 6 additional arguments
whose default is NULL:

```{bash}
# mdd
CHR SNP BP A1 A2 FRQ_A_59851 FRQ_U_113154 INFO OR SE P
awk '{print $2, $1, $3, $4, $5, $7, $8, $9, $10, $11}' MDD2018_ex23andMe.txt > mdd_2018.txt
# anxiety
# scz
awk '{print $2, $1, $3, $4, $5, $7, $8, $9, $10, $11}' scz_withRS.txt > schizophrenia.txt 
```

1.  files: The name of the summary statistics files. This should be the same as
    the name of the files used for the munge function in Step 1 above and the
    files should be in the same listed order used for the ldsc function in
    step 2.

```{r}
files = c(
   "GWAS_CP_all.txt",
   "cogent_g_n38.txt",
   "adhd_pgc_2019.txt",
   "anxiety.txt",
   "mdd_2018.txt",
   "scz_withRS.txt"
   )
```

1.  ref: The reference file used to calculate SNP variance across traits. We use
    1000 genomes phase 3 that is available for download here:
    <https://utexas.box.com/s/vkd36n197m8klbaio3yzoxsee6sxo11v>.

```{r}
ref = "reference.1000G.maf.0.005.txt"
```

1.  trait.names: The names of the traits in the order that they are listed for
    the files.

```{r}
trait.names <- c(
   "CP",
   "COG",
   "ADHD",
   "ANX",
   "MDD",
   "SCZ")
```

1.  se.logit: Whether the SEs are on a logistic scale. Please note that for many
    traits, the SEs will be reported on a logistic scale even when the
    regression effects are reported as odds ratios (ORs). This information can
    usually be found in the README file for the summary statistics. If this is
    listed as FALSE, and all other additional arguments are FALSE, then the
    function is expecting the standard error of the OR. The function will
    automatically determine whether the effect column is an OR or on a logistic
    scale based on the median effect size.

```{r}
se.logit = c(F,F,F,F,F,F)
```

1.  OLS: Whether the phenotype was a continuous outcome analyzed using an
    observed least square (OLS; i.e., linear) estimator. If no information is
    provided the default is for the function to assume FALSE. We write
    "OLS=NULL" for this argument, and all other arguments whose default is NULL,
    so it is clear what the argument names are when needed; however, this need
    not be written out if the argument is in fact FALSE for all of your traits.

```{r}
OLS = c(T,T,F,F,F,F)
```

1.  linprob: Whether the phenotype was a dichotomous outcome analyzed using an
    OLS estimator, as is the case for certain UKB phenotypes analyzed using the
    Hail software. In this case, a form of standardization is performed where
    the betas are converted to a logistic scale, and then standardized with
    respect to the outcome. If no information is provided the default is for the
    function to assume NULL.

```{r}
linprob = NULL
```

1.  prop: In order to perform the argument 6 conversion above from OLS betas to
    logistic betas, argument 7 takes the proportion of cases over the total
    sample size (range: 0 - 1). This argument should only contain numeric values
    if argument 6 is TRUE. If no information is provided the default is for the
    function to assume NULL.

```{r}
prop = NULL
```

1.  N: A user provided N listed in the order the traits are listed for the files
    argument. If no information is provided, the default is for the function to
    assume NULL. For LPM and OLS transformation outlined above the function uses
    sample size for conversions, so the summary statistics should contain a
    sample size (N) column or a sample size should be provided. If sample sizes
    are being provided for some traits, but not others, then NA can be used for
    traits that the user does not wish to provide sample sizes for. If the
    summary statistics file reports N separately for cases and controls, the
    function will automatically use both columns to compute total sample size.
    For this particular example, all traits are case/control so N is not used in
    the conversions and is not listed.

```{r}
N = c(
   257828,
   28445,
   55274,
   17310,
   173005,   
   105318)
```

1.  info.filter: The INFO filter to use. The package default is 0.6.

```{r}
info.filter = 0.6
```

1.  maf.filter: The MAF filter to use. The package default is 0.01.

```{r}
maf.filter = 0.01
```

1.  keep.indel: Whether insertion deletions (indels) should be included in the
    output. Default = FALSE.

```{r}
keep.indel = FALSE
```

1.  parallel: Whether the function should be run in parallel. Default = FALSE.

```{r}
parallel = FALSE # had to change this
```

1.  cores: When running in parallel, whether you want the computer to use a
    certain number of cores. If parallel is set to true, default is to use
    either 1 less than the total number of cores available, or when the number
    of available cores exceeds the number of traits to use as many cores as
    there are traits.

```{r}
cores = 8 # had to change this too 
```

### Run sumstats analysis

```{r}
# this is an R object, not a file
p_sumstats <- sumstats(
   files = files,
   ref = ref,
   trait.names = trait.names,
   se.logit = se.logit,
   OLS = OLS,
   linprob = linprob,
   prop = prop,
   N = N,
   info.filter = info.filter,
   maf.filter = maf.filter,
   keep.indel = keep.indel,
   parallel = parallel,
   cores = cores
   )

# save as file to look at it
readr::write_tsv(p_sumstats, "p_sumstats.txt")
```

## Step 4: Combine Summary Statistics and LDSC Output and Run the Common Factor GWAS

Note that as of 4/6/21, we combine the addSNPs and GWAS functions to save
memory. The output from addSNPs can no longer be fed to the GWAS functions. In
order to run multivariate GWAS, the S and V matrices from LDSC and the newly
prepared summary statistics files first need to be merged to create as many S
and V matrices as there are SNPs. The commonfactorGWAS function does this by
expanding the S and V matrices to include the SNP effects. In addition, the
function transforms the GWAS summary statistics (i.e., regression coefficients)
and their SEs into covariances and SEs of covariances by taking the product of
the regression coefficient and SNP variance from the reference file. Finally,
the function will smooth S and V matrices that are non-positive definite when
including the SNP covariances. After constructing the S and V matrices, the
function will run the common factor model.

Note that the commonfactorGWAS function use unit loading identification, such
that the loading of the first indicator is fixed to 1 so that the model is
identified. *In order for the model to converge, it will typically be best to
order your files in the ldsc and sumstats function such that the phenotype with
the largest loading on the common factor is listed first.* This can be
determined by running `commonfactor` on just the `ldsc matrix` to determine what
the factor structure looks like when not including the effect of an individual
SNP. In this case, the phenotype with the largest loading on the common factor
is schizophrenia. The `commonfactor` function has the following arguments:

```{r}
## this in-between step
covstruc = LDSCoutput
commonfactor(covstruc, estimation="DWLS")

## Result
>[1] "Running Model"
>[1] "Calculating CFI"
elapsed 
  0.304 
>$modelfit
   chisq df p_chisq AIC CFI        SRMR
df    NA  0      NA  NA   1 2.74995e-10

>$results
   lhs op  rhs Unstandardized_Estimate Unstandardized_SE Standardized_Est
1   F1 =~ ADHD              0.17902575        0.02168522       0.35284459
2   F1 =~  COG             -0.34838806        0.03318756      -0.93799971
3   F1 =~   CP             -0.46621474        0.04477092      -1.04419362
5 ADHD ~~ ADHD              0.22538271        0.01751785       0.87550070
6  COG ~~  COG              0.01657557        0.02636861       0.12015655
7   CP ~~   CP             -0.01800908        0.04123069      -0.09034031
  Standardized_SE      p_value
1      0.04273974 1.510699e-16
2      0.08935415 8.865107e-26
3      0.10027463 2.155288e-25
5      0.06804823 7.003479e-38
6      0.19114642 5.296044e-01
7      0.20682864 6.622649e-01

```

1.  covstruc: The output from LDSC.

```{r}
covstruc = LDSCoutput
```

1.  SNPs: The output from sumstats.

```{r}
SNPs = p_sumstats
```

1.  estimation: Whether the models should be estimated using "DWLS" or "ML"
    estimation. The package default is "DWLS".

```{r}
estimation = "DWLS"
```

1.  cores: How many computer cores to use in parallel processing. The default is
    to use 1 less core than is available in the local environment.

```{r}
cores = NULL
```

1.  toler: What tolerance level to use for matrix inversions. This is only
    something that needs to be of concern if warning/error messages are produced
    to the effect of "matrix could not be inverted".

```{r}
toler = FALSE
```

1.  SNPSE: An optional argument specifying whether you want to set the SE of the
    SNP to a value higher than the default of .0005. This number is set low to
    reflect that this value is essentially be treating as fixed.

```{r}
SNPSE = .0005
```

1.  parallel: An optional argument specifying whether you want the function to
    be run in parallel, or to be run serially. As most outputs from sumstats
    will be quite long (i.e., \> 1 million SNPs), running the function in
    parallel is highly recommended in order to save time; the default is for the
    function to run in parallel unless you are using a Windows operating system.

```{r}
parallel = TRUE
```

1.  GC: Level of Genomic Control (GC) you want the function to use. The default
    is 'standard' which adjusts the univariate GWAS standard errors by
    multiplying them by the square root of the univariate LDSC intercept.
    Additional options include 'conserv', which adjusts standard errors by
    multiplying with univariate LDSC intercept, and 'none' which does not
    correct the standard errors.

```{r}
GC = "standard"
```

1.  MPI: Whether the function should use multi-node processing (i.e., MPI).
    Please note that this should only be used on a computing cluster on which
    the R package Rmpi is already installed.

```{r}
MPI = FALSE
```

### Run CommonFactorGWAS

```{r}
# commonfactor ------------------------------------------------------------
covstruc <- LDSCoutput
cf <- GenomicSEM::commonfactor(covstruc, estimation = "DWLS")
```

```{r}
# run the multivariate GWAS using parallel processing
covstruc <- LDSCoutput
SNPs <- p_sumstats
estimation <- "DWLS"
cores <- NULL # change for CARC
toler <- FALSE
SNPSE <- FALSE
parallel <- FALSE # change for CARC
GC <- "standard"
MPI <- FALSE
# function
pfactor <- commonfactorGWAS(
    covstruc = covstruc, SNPs = SNPs, estimation = estimation,
    cores = cores, toler = toler, SNPSE = SNPSE, parallel = parallel,
    GC = GC, MPI = MPI
)
```

```{r}
# note that the code written above specifies all arguments for completeness, but as many of these arguments
# are set to the package default it could also be written as below and produce the same result:
pfactor <- commonfactorGWAS(covstruc = LDSCoutput, SNPs = p_sumstats)
```

```{r}
# to view submitted job
less slurm-<JOBID>.out

#pfactor <- commonfactorGWAS(
#   covstruc = LDSCoutput, 
#   SNPs = p_sumstats, 
#   estimation = "DWLS", 
#   cores = cores, 
#   toler = FALSE, 
#   SNPSE = FALSE, 
#   parallel = TRUE, 
#   GC = "standard",
#   MPI = FALSE
#)

# note that the code written above specifies all arguments for completeness, but as many of these arguments
# are set to the package default it could also be written as below and produce the same result:
pfactor <- commonfactorGWAS(
   covstruc = LDSCoutput,
   SNPs = p_sumstats)

```

The figure directly below depicts as a path diagram an example of the model run
for SNP rs4552973: <!-- insert image -->

The full set of summary statistics, along with a ReadMe file, can be found at:
<https://utexas.box.com/s/vkd36n197m8klbaio3yzoxsee6sxo11v> The first 6 columns
are pulled from the file created using sumstats and list the rsID, chromosome,
BP, MAF from the reference file, A1, and A2. The "i" column is simply an ordered
list of the runs. The "lhs", "op", and "rhs" column lists the outcome, operator,
and predictor. In this case "F1 \~ SNP", indicates the common factor regressed
on the SNP. The user should check that this column reads as such to confirm the
appropriate results are being pulled. The "est" column lists the effect of the
SNP on the common factor. The "se_c" column lists the sandwich corrected
standard error associated with this effect. The "Z_Estimate" column lists the
ratio of est/se_c. The "Pval_Estimate" column lists the corresponding p-value of
Z_Estimate.

The Q column lists the heterogeneity estimate for that SNP. This indexes the
extent to which the SNP effect is not mediated by the common factor, with larger
values indicating greater heterogeneity. This is a chi-square distributed test
statistic, with degrees of freedom for a common factor model equal to the number
of indicators - 1. The "Q_df" column lists the degrees of freedom for Q, and the
"Q_pval" column lists the p-value associated with the Q statistic. The fail
column indicates whether a particular run did not converge. The warning column
will list whether a particular run raised any warnings in lavaan. This may
include specific runs where the residual variance of an indicator is negative.

Calculating sample size for factors Once the summary statistics for a common
factor are produced, the user may wish to input those summary statistics into an
outside program that requires a sample size (e.g., LDSC, LDpred). A description
of one method for calculating effective sample size can be found in the online
supplement of bioRxiv pre-print from Mallard et al. (2019):
<https://www.biorxiv.org/content/10.1101/603134v1.abstract>. The equation
described in the supplement can be written as below to produce the effective
sample size for the p-factor. Note that the recommendation is also made to
restrict the summary statistics to MAF limits of approximately 10% and 40% in
order to produce more stable estimates.

```{r}
#restrict to MAF of 40% and 10%

pfactor2 <- subset(pfactor, pfactor$MAF <= .4 & pfactor$MAF >= .1)

#calculate effective N

Effective_N <-
   (mean(((
      pfactor2$Z_Estimate / pfactor2$est
   ) ^ 2) / (2 * pfactor2$MAF * (1 - pfactor2$MAF))))

```

We note also that for LDSC, the sample size provided does not effect the
univariate intercept or the intercept SE. Although the heritability estimate and
SE are contingent on user provided sample size, the ratio of the heritability
estimate over SE is retained regardless of the sample size provided and,
therefore, the heritability z-statistic can still be calculated.

The commonfactorGWAS and commonfactorGWASpar functions use unit loading
identification and genetic covariance matrices (and corresponding sampling
covariance matrices) as input. That is, the SNP effects shown in the figure and
table above are unstandardized with respect to both the outcome (the common
factor) and the predictor (the SNP). As the common factor is a dependent
variable, its variance is not directly represented, but rather implied. Thus,
unit variance identification would require a model constraint. The latest
version includes a user defined GWAS function in which the the user has
substantial flexibility regarding what covariance structure and parameterization
(such as specifying a model constraint) is specified. Output for a fully
standardized version of a common factor GWAS is shown as an example below for
rs4552973:

A note on serial versus parallel processing. Serial and parallel processing will
return the same output, but are computationally implemented in a serial and
parallel fashion, respectively. In our experience using high performance
computing systems there were often memory issues when using parallel processing
unless you have access to large memory nodes. If memory issues are a problem
then serial processing may be preferred. However, on a personal laptop with
multiple cores, the computing time will be reduced greatly when using parallel
processing. On a personal laptop with 8 cores using parallel processing for a
common factor with five indicators and 1 million SNPs, the run time was
approximately 8 hours. These functions are running as many models as there are
SNPs.

# 3. Models without Individual SNP effects

## Step 1: Munge the summary statistics

```{r}
##read in the schizophrenia summary statistics using fread from the data.table R package##
require(data.table)
scz<-fread("clozuk_pgc2.meta.sumstats.txt",data.table=FALSE) 
##pull out the rsIDs##
scz$SNP<-sapply(strsplit(scz$SNP, ":"), `[`, 1)
##output the result in a .txt file called scz_withRS.txt##
write.table(scz, file = "scz_withRS.txt", sep = "\t", quote = FALSE, row.names = FALSE, col.names=TRUE)
##munge the summary statistics files##
##note that we include the argument names (e.g., files, hm3) for
##completeness in the code below, but that this is not necessary to run the code.
munge(files=c("scz_withRS.txt", "pgc.bip.full.2012-04.txt", "MDD2018_ex23andMe.txt", "SORTED_PTSD_EA9_ALL_study_specific_PCs1.txt", "anxiety.meta.full.cc.tbl"), hm3 = "w_hm3.snplist",trait.names=c("scz", "bip", "mdd", "ptsd", "anx"),N=c(105318, 16731, 173005, 9537, 17310), info.filter = 0.9, maf.filter = 0.01)
```

```{r}
#GWM
scz<-fread("clozuk_pgc2.meta.sumstats.txt",data.table=FALSE) 

```

## Step 2: Run multivariable LDSC

```{r}
traits <- c("scz.sumstats.gz", "bip.sumstats.gz", "mdd.sumstats.gz", "ptsd.sumstats.gz", "anx.sumstats.gz")
sample.prev <- c(.39,.45,.35,.25,.32)
population.prev <- c(.01,.01,.16,.08,.29)
ld <- "eur_w_ld_chr/"
wld <- "eur_w_ld_chr/"
trait.names<-c("SCZ","BIP","MDD","PTSD","ANX")
LDSCoutput <- ldsc(traits, sample.prev, population.prev, ld, wld, trait.names)

##optional command to save the ldsc output in case you want to use it in a later R session. 
save(LDSCoutput, file="Pfactor.RData")

k<-nrow(LDSCoutput$S)
SE<-matrix(0, k, k)
SE[lower.tri(SE,diag=TRUE)] <-sqrt(diag(LDSCoutput$V))


```

## Common Factor Model

```{r}
#To run using DWLS estimation#
CommonFactor_DWLS<- commonfactor(covstruc = LDSCoutput, estimation="DWLS")

#print CommonFactor_DWLs output#
CommonFactor_DWLS

```

```{r}
commonfactor.model<-'F1=~ NA*SCZ + BIP + MDD + PTSD + ANX
F1~~1*F1
```

## User-Specified Models

Genomic SEM is not limited to common factor models, and can be used to fit any
form of structural model, enabling the ability to test a myriad of models and
theoretical accounts. We outline some examples of how to specify your own models
here, starting with the GWIS example from the paper.

### Genetic multiple regression (similar to GWIS)

```{r}
#munge the files
munge(c("pgc.cross.SCZ17.2013-05.txt", "pgc.cross.BIP11.2013-05.txt", "SSGAC_EduYears_Rietveld2013_publicrelease.txt"), hm3 = "w_hm3.snplist",trait.names=c("sczcross", "bipcross", "EA"), c(26552, 28675, 126559), info.filter = 0.9, maf.filter = 0.01)

#run multivariable LDSC to create the S and V matrices
ld <- "eur_w_ld_chr/"
wld <- "eur_w_ld_chr/"
traits <- c("sczcross.sumstats.gz","bipcross.sumstats.gz", "EA.sumstats.gz")
sample.prev <- c(.27,.32,NA)
population.prev <- c(.01,.01,NA)
trait.names<-c("SCZ", "BIP", "EA")

GWIS<-ldsc(traits, sample.prev, population.prev, ld, wld, trait.names)

```

```{r}
#specify a model in which Educational Achievement is regressed on Bipolar Disorder and Schizophrenia, and the correlation is modeled #between Bipolar and Schizophrenia
GWISmodel <- 'EA ~ SCZ + BIP
SCZ~~BIP'

#run the model using the usermodel function
GWISoutput<-usermodel(covstruc = GWIS, estimation = "DWLS", model = GWISmodel, CFIcalc = TRUE, std.lv = FALSE, imp_cov = FALSE)

#print the output
GWISoutput
```

## Exploratory Factor Analysis

```{r}
#munge the summary statistics
munge(c("SNP_gwas_mc_merge_nogc.tbl.uniq", "GIANT_2015_WHR_COMBINED_EUR.txt", "EGG_Obesity_Meta_Analysis_1.txt", "GIANT_2015_HIP_COMBINED_EUR.txt", "GIANT_2015_WC_COMBINED_EUR.txt", "GIANT_HEIGHT_Wood_et_al_2014_publicrelease_HapMapCeuFreq.txt", "EGG_HC_DISCOVERY.v2.txt", "EGG-GWAS-BL.txt", "EGG_BW2_DISCOVERY.txt"), hm3 = "w_hm3.snplist",trait.names=c("BMI2015", "waisthip", "childobese", "hip", "waist", "height", "headcirc", "birthlength", "birthweight"), c(NA, NA, 13848, NA, NA, NA, NA, NA, 26836), info.filter = 0.9, maf.filter = 0.01)

#run multivariable LDSC to create the S and V matrices
ld <- "eur_w_ld_chr/"
wld <- "eur_w_ld_chr/"
traits<-c("BMI2015.sumstats.gz", "waisthip.sumstats.gz", "childobese.sumstats.gz", "waist.sumstats.gz","hip.sumstats.gz", "height.sumstats.gz", "headcirc.sumstats.gz", "birthlength.sumstats.gz", "birthweight.sumstats.gz")
sample.prev <- c(NA,NA,NA,NA,NA,NA,NA,NA,NA)
population.prev <- c(NA,NA,NA,NA,NA,NA,NA,NA,NA)
trait.names<-c("BMI","WHR","CO","Waist", "Hip", "Height", "IHC", "BL", "BW")
anthro<-ldsc(traits, sample.prev, population.prev, ld, wld, trait.names)

#smooth the S matrix for EFA using the nearPD function in the Matrix package. 
require(Matrix)
Ssmooth<-as.matrix((nearPD(anthro$S, corr = FALSE))$mat)

#run EFA with promax rotation and 2 factors using the factanal function in the stats package
require(stats)
EFA<-factanal(covmat = Ssmooth, factors = 2, rotation = "promax")

#print the loadings
EFA$loadings

```

## Confirmatory factor analysis

```{r}
#Specify the Genomic confirmatory factor model
CFAofEFA <- 'F1 =~ NA*BMI + WHR + CO + Waist + Hip
             F2 =~ NA*Hip + Height + IHC + BL + BW
F1~~F2
Waist ~~ a*Waist
a > .001'

#run the model
Anthro<-usermodel(anthro, estimation = "DWLS", model = CFAofEFA, CFIcalc = TRUE, std.lv = TRUE, imp_cov = FALSE)

#print the results
Anthro
```
